---
output: word_document
author: "GH/OHA/SPS/HWF"
date: "`r format(Sys.time(), '%d %B, %Y')`"
params: 
    set_title: "My Title!"
    operating_unit: !r c("Tanzania")
---

---
title: "FY23 HRH Inventory Briefer - `r params$operating_unit`"
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE) # supress code
knitr::opts_chunk$set(warning = FALSE, message = FALSE) # suppress all warnings
knitr::opts_chunk$set(dpi=300) # set high resolution of plots

library(plyr)
library(dplyr)
library(tidyverse)
library(sqldf)
library(janitor)
library(readxl)
library(writexl)
library(shiny)
library(scales)
library(tidyquant)
library(formattable)
library(kableExtra)

```

## Automated reports at OU level

This provides a detailed summary of the FY23 HRH Inventory dataset for `r params$operating_unit`. These reports help summarize staffing investments and trends at the country level to inform current programming, and facilitate HRH planning.

```{r}

# Load the FSD and HRH dataset
fin_data_orig <- read.delim("./1. Data/Financial_Structured_Datasets_COP17-23_20231215.txt") 
load(file = "./4. Outputs/RDS/FY23_cleanHRH.rds") 
load(file = "./4. Outputs/RDS/HRH_ER_merged_21_23.rds") 

```

## USAID and CDC HRH Inventory Submission Rate
This shows the number of mechanisms that submitted both ER and HRH templates to better understand the overall completeness/reliability of the HRH inventory.

```{r, echo = FALSE, eval = TRUE}


mech_submissions <- finalMerge %>%
  filter(year == 2023) %>%
  filter(GHSC_UN_keywordflags == "FALSE") %>% # excluding all GHSC and UN mechanisms
  filter(operating_unit == params$operating_unit) %>%
  group_by(year, operating_unit, funding_agency, mech_code) %>%
  summarise(HRH_expenditure_amt = sum(HRH_expenditure_amt, na.rm = T), 
            ER_expenditure_amt = sum(ER_expenditure_amt[HRH_relevant == "Y"], na.rm = T)) %>%
  ungroup() %>%
  group_by(funding_agency) %>%
  summarise(Num_HRH_submissions = length(mech_code[HRH_expenditure_amt > 0]),
            Num_ER_submissions = length(mech_code[ER_expenditure_amt > 0]),
            HRH_expenditure_amt = sum(HRH_expenditure_amt, na.rm = T),
            ER_expenditure_amt = sum(ER_expenditure_amt, na.rm = T)) %>%
  ungroup() %>%
  arrange(desc(HRH_expenditure_amt)) %>%
  adorn_totals() %>%
  mutate(pct_submission_rate = HRH_expenditure_amt / ER_expenditure_amt) %>%
  
  # format the columns
  mutate(across(4:5, formattable::currency,digits=0)) %>%
  mutate(across(6, formattable::percent, digits=0))

# Now, display the final table with formatted column names
knitr::kable(mech_submissions, "simple", 
             col.names=c("Funding Agency", "No. of HRH Submissions", "No. of ER Submissions", 
                         "HRH Expenditure Amount (USD)", "ER Expenditure Amount   (USD)", "Overall Submission Rate (%)"),
             caption = "Submission Rate by Agency")


```

## Initial Summary of PEPFAR vs USAID HRH Staffing footprint
This table provides a summary of USAID's staffing size relative to PEPFAR in `r params$operating_unit`. Relative to PEPFAR, this shows USAID's relative staffing footprint (individual count and FTE), the % going towards staffing, and the amount spent by local prime implementation partners

```{r echo = FALSE, eval = TRUE}

# Create the first df that summarizes PEPFAR
infographic_PEPFAR <- HRH_clean %>%
  filter(fiscal_year == 2023) %>%
  filter(operating_unit == params$operating_unit) %>%
  mutate(funding_agency = "PEPFAR") %>%
  group_by(funding_agency) %>%
  summarise(HRH_expenditure = sum(actual_annual_spend, na.rm = T),
            individual_count = sum(individual_count, na.rm = T),
            annual_fte = sum(annual_fte, na.rm = T),
            placeholder_pctHRH = sum(actual_annual_spend, na.rm = T),
            local_partner_expenditures = sum(actual_annual_spend[is_indigenous_prime_partner == "Y"]))

# Calculate the % towards staffing
PEPFAR_pct_HRH <- finalMerge %>%
  filter(year == 2023) %>%
  filter(operating_unit == params$operating_unit) %>%
  group_by(year, operating_unit) %>%
  summarise(HRH_expenditure_amt = sum(HRH_expenditure_amt, na.rm = T),
            ER_expenditure_amt = sum(ER_expenditure_amt, na.rm = T)) %>%
  ungroup() %>%
  mutate(PEPFAR_pct_HRH = HRH_expenditure_amt / ER_expenditure_amt) %>%
  pull(PEPFAR_pct_HRH)

# Replace the % HRH placeholder value with the calculated value
infographic_PEPFAR <- infographic_PEPFAR %>%
  mutate(placeholder_pctHRH = PEPFAR_pct_HRH) %>%
  rename(pct_HRH = placeholder_pctHRH)

# Now, create the second df that summarizes USAID only
infographic_USAID <- HRH_clean %>%
  filter(fiscal_year == 2023) %>%
  filter(operating_unit == params$operating_unit) %>%
  filter(funding_agency == "USAID") %>%
  group_by(funding_agency) %>%
  summarise(HRH_expenditure = sum(actual_annual_spend, na.rm = T),
            individual_count = sum(individual_count, na.rm = T),
            annual_fte = sum(annual_fte, na.rm = T),
            placeholder_pctHRH = sum(actual_annual_spend, na.rm = T),
            local_partner_expenditures = sum(actual_annual_spend[is_indigenous_prime_partner == "Y"]))

# Calculate the % towards staffing for USAID
USAID_pct_HRH <- finalMerge %>%
  filter(year == 2023) %>%
  filter(operating_unit == params$operating_unit) %>%
  filter(funding_agency == "USAID") %>%
  group_by(year, operating_unit) %>%
  summarise(HRH_expenditure_amt = sum(HRH_expenditure_amt, na.rm = T),
            ER_expenditure_amt = sum(ER_expenditure_amt, na.rm = T)) %>%
  ungroup() %>%
  mutate(USAID_pct_HRH = HRH_expenditure_amt / ER_expenditure_amt) %>%
  pull(USAID_pct_HRH)

# Replace the % HRH placeholder value with the calculated value for USAID
infographic_USAID <- infographic_USAID %>%
  mutate(placeholder_pctHRH = USAID_pct_HRH) %>%
  rename(pct_HRH = placeholder_pctHRH)

# Lastly, bind both data frames together to create the final summary table
infographic <- rbind(infographic_PEPFAR, infographic_USAID) %>%


# format the columns
  mutate(across(2, formattable::currency, digits=0)) %>%
  mutate(across(3:4, formattable::accounting, digits=0)) %>%
  mutate(across(5, formattable::percent, digits=0)) %>%
  mutate(across(6, formattable::currency, digits=0))

# Display the final summary table with formatted column names
knitr::kable(infographic, "simple", 
             col.names=c("Funding Agency", "HRH Staffing Expenditure", "Individual Count", 
                         "Total FTE (Full-Time Equivalence)", "% of Total Expenditures Spent Towards HRH", 
                         "Local Partner HRH Staffing Expenditure"),
             caption = "FY23 HRH Summary: By The Numbers")

```



## FY23 Staffing Footprint by Funding Agency
Breakdown of HRH expenditure and number of individual staff supported by each agency

```{r, echo = FALSE, eval = TRUE}

## Create a Transformer Function to be used by the rest of the dual axis plots ----
transformer_dual_y_axis <- function(data,
                                    primary_column, secondary_column,
                                    include_y_zero = FALSE) {

    # PARAMETER SETUP
    params_tbl <- data %>%
        summarise(
            max_primary   = max(!! enquo(primary_column)),
            min_primary   = min(!! enquo(primary_column)),
            max_secondary = max(!! enquo(secondary_column)),
            min_secondary = min(!! enquo(secondary_column))
        )

    if (include_y_zero) {
        params_tbl$min_primary   <- 0
        params_tbl$min_secondary <- 0
    }

    params_tbl <- params_tbl %>%
        mutate(
            scale = (max_secondary - min_secondary) / (max_primary - min_primary),
            shift = min_primary - min_secondary
        )

    # MAKE SCALER FUNCTIONS
    scale_func <- function(x) {
        x * params_tbl$scale - params_tbl$shift
    }

    inv_func <- function(x) {
        (x + params_tbl$shift) / params_tbl$scale
    }

    # RETURN
    ret <- list(
        scale_func = scale_func,
        inv_func   = inv_func,
        params_tbl = params_tbl
    )

    return(ret)
}

```

```{r fig1, fig.height = 4, fig.width = 7, echo = FALSE, eval = TRUE}

# Create a df that shows HRH expenditure and individual count by program area
dual_axis <- HRH_clean %>%
  filter(fiscal_year == 2023) %>%
  filter(operating_unit == params$operating_unit) %>%
  select(fiscal_year, funding_agency, actual_annual_spend, individual_count) %>%
  group_by(funding_agency) %>%
  summarise(actual_annual_spend = sum(actual_annual_spend, na.rm = T),
            individual_count = sum(individual_count, na.rm = T)) %>%
  arrange(desc(individual_count)) 

# Make the Y-Axis Transformer to properly scale / align the dual axis in the plot
transformer <- dual_axis %>%
    transformer_dual_y_axis(
        primary_column   = individual_count,
        secondary_column = actual_annual_spend,
        include_y_zero   = TRUE
    )

# Now make the plot

g1 <- dual_axis %>%
    ggplot(aes(x = reorder(funding_agency, -individual_count))) +
    geom_col(
        aes(y = individual_count, fill = "Individual Count"),
        alpha = 0.9,
        width = 0.7
    ) +
    geom_label(
        aes(
            y     = individual_count,
            label = scales::comma(round(individual_count)),
            color = "Individual Count",
        ),
        size = 3
    )

# * 3.2 SECONDARY AXIS ----

# Create a y-label distance for use in this graph based on the max value of the y-axis
y_axis_nudge <- dual_axis %>%
  summarise(max_y_value = max(actual_annual_spend)) %>%
  mutate(nudge = max_y_value * .000017) %>%
  pull(nudge)

g2 <- g1 +
    geom_point(
        aes(y     = transformer$inv_func(actual_annual_spend),
            color = "HRH Expenditure"),
        size = 4
    ) +
    geom_label(
        aes(
            y     = transformer$inv_func(actual_annual_spend),
            label = scales::dollar(round(actual_annual_spend)),
            color = "HRH Expenditure"),
        size = 3,
        nudge_y = y_axis_nudge
    ) + 
  
    scale_y_continuous(
        name     = "Individual Count",
        labels = scales::comma,
        sec.axis = sec_axis(
            trans = ~ transformer$scale_func(.),
            name  = "Total HRH Expenditure",
            labels = scales::dollar
        )) +
  
  labs(title = stringr::str_glue("Individual Count and HRH Expenditure by Funding Agency in {params$operating_unit}")) +
  
  theme(panel.grid = element_blank(), panel.background = element_rect(fill = "white"),
        axis.line=element_blank(),
        axis.text.y=element_blank(),
        axis.ticks=element_blank(),
        axis.title.x=element_blank(),
        axis.title.y=element_blank(),
        plot.title = element_text(size = 10)) 

funding_agency_summary <- g2 +
    theme_tq() +
    scale_color_manual(values = c(
        palette_light()[["red"]],
        palette_light()[["blue"]]
    )) +
    scale_fill_manual(values = palette_light()[["blue"]]) +
    theme(
        legend.position    = "none",
        axis.title.y.right = element_text(color = palette_light()[["red"]]),
        axis.text.y.right  = element_text(color = palette_light()[["red"]]),
        axis.title.x=element_blank()
    )

funding_agency_summary

```



## FY23 Staffing Breakdown by Program Area
Breakdown of total HRH expenditure and number of individual staff by program area

```{r fig2, fig.height = 4.5, fig.width = 8, echo = FALSE, eval = TRUE}

# Create a df that shows HRH expenditure and individual count by program area
dual_axis <- HRH_clean %>%
  filter(fiscal_year == 2023) %>%
  filter(funding_agency == "USAID") %>%
  filter(operating_unit == params$operating_unit) %>%
  filter(!program == "Error") %>%
  select(fiscal_year, funding_agency, program, actual_annual_spend, individual_count) %>%
  group_by(program) %>%
  summarise(actual_annual_spend = sum(actual_annual_spend, na.rm = T),
            individual_count = sum(individual_count, na.rm = T)) %>%
  arrange(desc(individual_count)) 

# Create % of total table for inline text later
dual_axis_pct_PM <- dual_axis %>%
  adorn_percentages("col") %>%
  mutate(actual_annual_spend = round(actual_annual_spend * 100),
         individual_count = round(individual_count * 100)) %>%
  filter(program == "PM") %>%
  pull(actual_annual_spend)

# Make A Y-Axis Transformer
transformer <- dual_axis %>%
    transformer_dual_y_axis(
        primary_column   = individual_count,
        secondary_column = actual_annual_spend,
        include_y_zero   = TRUE
    )

# Make the first plot

g1 <- dual_axis %>%
    ggplot(aes(x = reorder(program, -individual_count))) +
    geom_col(
        aes(y = individual_count, fill = "Individual Count"),
        alpha = 0.9,
        width = 0.8
    ) +
    geom_label(
        aes(
            y     = individual_count,
            label = scales::comma(round(individual_count)),
            color = "Individual Count",
        ),
        size = 3
    )

# * 3.2 SECONDARY AXIS ----

# Create a y-label distance for use in this graph based on the max value of the y-axis
y_axis_nudge <- dual_axis %>%
  summarise(max_y_value = max(actual_annual_spend)) %>%
  mutate(nudge = max_y_value * .000026) %>%
  pull(nudge)

g2 <- g1 +
    geom_point(
        aes(y     = transformer$inv_func(actual_annual_spend),
            color = "HRH Expenditure"),
        size = 4
    ) +
    geom_label(
        aes(
            y     = transformer$inv_func(actual_annual_spend),
            label = scales::dollar(round(actual_annual_spend)),
            color = "HRH Expenditure"),
        size = 3,
        nudge_y = y_axis_nudge
    ) + 
  
  
#scale_y_continuous(labels = scales::comma)
    scale_y_continuous(
        name     = "Individual Count",
        labels = scales::comma,
        sec.axis = sec_axis(
            trans = ~ transformer$scale_func(.),
            name  = "Total HRH Expenditure",
            labels = scales::dollar
        )
    ) +
  
  labs(title = stringr::str_glue("Individual Count and HRH Expenditure by Program Area in {params$operating_unit}")) +
  
  theme(panel.grid = element_blank(), panel.background = element_rect(fill = "white"),
        axis.line=element_blank(),
        axis.text.y=element_blank(),
        axis.ticks=element_blank(),
        axis.title.x=element_blank(),
        axis.title.y=element_blank(),
        plot.title = element_text(size = 10)) 

g3 <- g2 +
    theme_tq() +
    scale_color_manual(values = c(
        palette_light()[["red"]],
        palette_light()[["blue"]]
    )) +
    scale_fill_manual(values = palette_light()[["blue"]]) +
    theme(
        legend.position    = "none",
        axis.title.y.right = element_text(color = palette_light()[["red"]]),
        axis.text.y.right  = element_text(color = palette_light()[["red"]]),
        axis.title.x=element_blank()
    )

g3

```

## FY23 Top Employment Titles (by count)
Breakdown of total HRH expenditure and number of individual staff by highest count

```{r fig3, fig.height = 5, fig.width = 8, echo = FALSE, eval = TRUE}

# Create a df that shows HRH expenditure and individual count by the top 10 employment title
dual_axis <- HRH_clean %>%
  filter(fiscal_year == 2023) %>%
  filter(funding_agency == "USAID") %>%
  filter(operating_unit == params$operating_unit) %>%
  filter(!program == "Error") %>%
  select(fiscal_year, funding_agency, employment_title, actual_annual_spend, individual_count) %>%
  group_by(employment_title) %>%
  summarise(actual_annual_spend = sum(actual_annual_spend, na.rm = T),
            individual_count = sum(individual_count, na.rm = T)) %>%
  arrange(desc(individual_count)) %>%
  top_n(n = 5, wt = individual_count) %>%
  ungroup() %>%
  mutate(employment_title_wrapped = str_wrap(employment_title, width = 10)) %>%
  select(-employment_title)

# Make A Y-Axis Transformer
transformer <- dual_axis %>%
    transformer_dual_y_axis(
        primary_column   = individual_count,
        secondary_column = actual_annual_spend,
        include_y_zero   = TRUE
    )

# Make the first plot
g1 <- dual_axis %>%
    ggplot(aes(x = reorder(employment_title_wrapped, -individual_count))) +
    geom_col(
        aes(y = individual_count, fill = "Individual Count"),
        alpha = 0.9,
        width = 0.8
    ) +
    geom_label(
        aes(
            y     = individual_count,
            label = scales::comma(round(individual_count)),
            color = "Individual Count",
        ),
        size = 3
    )

# * 3.2 SECONDARY AXIS ----

# Create a y-label distance for use in this graph based on the max value of the y-axis
y_axis_nudge <- dual_axis %>%
  summarise(max_y_value = max(actual_annual_spend)) %>%
  mutate(nudge = max_y_value * .00017) %>%
  pull(nudge)

g2 <- g1 +
    geom_point(
        aes(y     = transformer$inv_func(actual_annual_spend),
            color = "HRH Expenditure"),
        size = 4
    ) +
    geom_label(
        aes(
            y     = transformer$inv_func(actual_annual_spend),
            label = scales::dollar(round(actual_annual_spend)),
            color = "HRH Expenditure"),
        size = 3,
        nudge_y = y_axis_nudge
    ) + 
  
  
#scale_y_continuous(labels = scales::comma)
    scale_y_continuous(
        name     = "Individual Count",
        labels = scales::comma,
        sec.axis = sec_axis(
            trans = ~ transformer$scale_func(.),
            name  = "Total HRH Expenditure",
            labels = scales::dollar
        )
    ) +
  
  labs(title = stringr::str_glue("Individual Count and HRH Expenditure by Top Employment Title (for count)")) +
  
  theme(panel.grid = element_blank(), panel.background = element_rect(fill = "white"),
        axis.line=element_blank(),
        axis.text.y=element_blank(),
        axis.ticks=element_blank(),
        axis.title.x=element_blank(),
        axis.title.y=element_blank(),
        plot.title = element_text(size = 10)) 

g3 <- g2 +
    theme_tq() +
    scale_color_manual(values = c(
        palette_light()[["red"]],
        palette_light()[["blue"]]
    )) +
    scale_fill_manual(values = palette_light()[["blue"]]) +
    theme(
        legend.position    = "none",
        axis.title.y.right = element_text(color = palette_light()[["red"]]),
        axis.text.y.right  = element_text(color = palette_light()[["red"]]),
        axis.title.x=element_blank()
    )

g3

```
  
## FY23 Top Employment Titles (by expenditure)
Breakdown of total HRH expenditure and number of individual staff by the highest expenditure

```{r fig4, fig.height = 5, fig.width = 8, echo = FALSE, eval = TRUE}

# Create a df that shows HRH expenditure and individual count by the top 10 employment title
dual_axis <- HRH_clean %>%
  filter(fiscal_year == 2023) %>%
  filter(funding_agency == "USAID") %>%
  filter(operating_unit == params$operating_unit) %>%
  filter(!program == "Error") %>%
  select(fiscal_year, funding_agency, employment_title, actual_annual_spend, individual_count) %>%
  group_by(employment_title) %>%
  summarise(actual_annual_spend = sum(actual_annual_spend, na.rm = T),
            individual_count = sum(individual_count, na.rm = T)) %>%
  arrange(desc(actual_annual_spend)) %>%
  top_n(n = 5, wt = actual_annual_spend) %>%
  ungroup() %>%
  mutate(employment_title_wrapped = str_wrap(employment_title, width = 10)) %>%
  select(-employment_title)

# Make A Y-Axis Transformer
transformer <- dual_axis %>%
    transformer_dual_y_axis(
        primary_column   = individual_count,
        secondary_column = actual_annual_spend,
        include_y_zero   = TRUE
    )

# Make the first plot
g1 <- dual_axis %>%
    ggplot(aes(x = reorder(employment_title_wrapped, -actual_annual_spend))) +
    geom_col(
        aes(y = individual_count, fill = "Individual Count"),
        alpha = 0.9,
        width = 0.8
    ) +
    geom_label(
        aes(
            y     = individual_count,
            label = scales::comma(round(individual_count)),
            color = "Individual Count",
        ),
        size = 3
    )

# * 3.2 SECONDARY AXIS ----

# Create a y-label distance for use in this graph based on the max value of the y-axis
y_axis_nudge <- dual_axis %>%
  summarise(max_y_value = max(actual_annual_spend)) %>%
  mutate(nudge = max_y_value * .00017) %>%
  pull(nudge)

g2 <- g1 +
    geom_point(
        aes(y     = transformer$inv_func(actual_annual_spend),
            color = "HRH Expenditure"),
        size = 4
    ) +
    geom_label(
        aes(
            y     = transformer$inv_func(actual_annual_spend),
            label = scales::dollar(round(actual_annual_spend)),
            color = "HRH Expenditure"),
        size = 3,
        nudge_y = y_axis_nudge
    ) + 
  
  
    scale_y_continuous(
        name     = "Individual Count",
        labels = scales::comma,
        sec.axis = sec_axis(
            trans = ~ transformer$scale_func(.),
            name  = "Total HRH Expenditure",
            labels = scales::dollar
        )
    ) +
  
  labs(title = stringr::str_glue("Individual Count and HRH Expenditure by Top Employment Title (for expenditure)")) +
  
  theme(panel.grid = element_blank(), panel.background = element_rect(fill = "white"),
        axis.line=element_blank(),
        axis.text.y=element_blank(),
        axis.ticks=element_blank(),
        axis.title.x=element_blank(),
        axis.title.y=element_blank(),
        plot.title = element_text(size = 10)) 

g3 <- g2 +
    theme_tq() +
    scale_color_manual(values = c(
        palette_light()[["red"]],
        palette_light()[["blue"]]
    )) +
    scale_fill_manual(values = palette_light()[["blue"]]) +
    theme(
        legend.position    = "none",
        axis.title.y.right = element_text(color = palette_light()[["red"]]),
        axis.text.y.right  = element_text(color = palette_light()[["red"]]),
        axis.title.x=element_blank()
    )

g3

```

## FY23 Median Annual Remuneration
This box plot shows the range of annual remuneration (i.e. annual salary) of the top employment titles


```{r fig5, fig.height = 7, fig.width = 9, echo = FALSE, eval = TRUE}

# Prepare the remuneration data frame
remuneration <- HRH_clean %>%
  filter(fiscal_year == 2023) %>%
  filter(funding_agency == "USAID") %>%
  filter(operating_unit == params$operating_unit) %>%
  filter(!program == "Error") %>%
  select(employment_title, equivalent_annual_salary)

# Remove any outliers for now
detect_outlier <- function(x) {
  # calculate first quantile
  Quantile1 <- quantile(x, probs=.25)
  # calculate third quantile
  Quantile3 <- quantile(x, probs=.90)
  # calculate inter quartile range
  IQR = Quantile3-Quantile1
  # return true or false
  x > Quantile3 + (IQR*1.5) | x < Quantile1 - (IQR*1.5)
}
 
# create remove outlier function
remove_outlier <- function(dataframe,
                            columns=names(dataframe)) {
  # for loop to traverse in columns vector
  for (col in columns) {
    # remove observation if it satisfies outlier function
    dataframe <- dataframe[!detect_outlier(dataframe[[col]]), ]
  }
  
  return(dataframe)
}


remuneration <- remove_outlier(remuneration, 'equivalent_annual_salary')

# Calculate the median annual remuneration by employment title, then filter for top 10 titles, pull the titles and use this as the sort variable
sort_list <- remuneration %>%
  group_by(employment_title) %>%
  summarise(median_remuneration = median(equivalent_annual_salary, na.rm = T),
            n = n()) %>%
  arrange(desc(median_remuneration)) %>%
  filter(n > 10) %>%
  top_n(n = 10, wt = median_remuneration) %>%
  mutate(employment_title_wrapped = str_wrap(employment_title, width = 10)) %>%
  pull(employment_title)
  
remuneration <- remuneration %>%
  filter(employment_title %in% sort_list)

# Create the box plot
boxplot2 <- ggplot(remuneration, aes(x=reorder(employment_title, equivalent_annual_salary, FUN = median), y=equivalent_annual_salary)) + 
  geom_boxplot(fill = NA, coef = 0)

boxplot2 + coord_flip() + labs(x = "Employment Title", y = "Median Annual Remuneration") + geom_jitter(width = 0.3, fill = 'lightblue')

# NEXT STEPS: Complete the rest of the bar graphs, including stacked bar graphs!!

```

## FY23 Local Partner Staffing Expenditures by Program Area

```{r fig6, fig.height = 5, fig.width = 8, echo = FALSE, eval = TRUE}

# Create local partners summary table

local_partner <- HRH_clean %>%
  filter(fiscal_year == 2023) %>%
  filter(funding_agency == "USAID") %>%
  filter(operating_unit == params$operating_unit) %>%
  filter(!program == "Error") %>%
  select(fiscal_year, funding_agency, program, is_indigenous_prime_partner, actual_annual_spend) %>%
  group_by(program, is_indigenous_prime_partner) %>%
  summarise(actual_annual_spend = sum(actual_annual_spend, na.rm = T)) %>%
  arrange(desc(actual_annual_spend)) 

ggplot(local_partner, aes(x = reorder(program, -actual_annual_spend), y = actual_annual_spend, fill = is_indigenous_prime_partner)) +
  geom_col() +
  
  geom_label(
        aes(
            y     = actual_annual_spend,
            label = scales::dollar(round(actual_annual_spend))),
        size = 3
    ) +
  
  theme(panel.grid = element_blank(), panel.background = element_rect(fill = "white"),
        axis.line=element_blank(),
        axis.text.y=element_blank(),
        axis.ticks=element_blank(),
        axis.title.x=element_blank(),
        axis.title.y=element_blank(),
        plot.title = element_text(size = 10)) +
  
  scale_fill_brewer(palette = "Pastel1") 

#NEXT STEP: Just move on to the next graphs, we will deal with formatting and polishing later

```





