---
output: word_document
author: "GH/OHA/SPS/HWF"
date: "`r format(Sys.time(), '%d %B, %Y')`"
params: 
    set_title: "My Title!"
    operating_unit: !r c("Tanzania")
    fiscal_year: "FY24"
---

---
title: "`r params$fiscal_year` Human Resources for Health (HRH) Inventory Summary - `r params$operating_unit`"
---

```{r setup, include=FALSE}

# set global standards
knitr::opts_chunk$set(echo = FALSE) # supress code outputs
knitr::opts_chunk$set(warning = FALSE, message = FALSE) # suppress all warnings
knitr::opts_chunk$set(dpi=300) # set high resolution of plots

# load all needed libraries
library(plyr)
library(dplyr)
library(tidyverse)
library(sqldf)
library(janitor)
library(readxl)
library(writexl)
library(shiny)
library(scales)
library(tidyquant)
library(formattable)
library(kableExtra)
library(ggrepel)
library(ggplot2)
library(patchwork)

```

## Background

This HRH Inventory Briefer provides a detailed summary of the `r params$fiscal_year` HRH Inventory for `r params$operating_unit`. This report will summarize the submitted HRH inventory templates by prime implementation partners to help country teams better monitor and understand their staffing footprint. 

Mission teams should use these briefers to review their ongoing staffing investments/trends at the OU level to inform current programming, and facilitate HRH planning. If there are questions on the summary info provided here, please reach out to iborces@usaid.gov

```{r}

# Load the FSD, HRH, and merged FSD-HRH dataset
fin_data_orig <- read.delim("./1. Data/Financial_Structured_Datasets_COP17-23_20231215.txt") # FSD
load(file = "./4. Outputs/RDS/FY23_cleanHRH.rds") # HRH
load(file = "./4. Outputs/RDS/HRH_ER_merged_21_23.rds") # FSD-HRH 

# Some data cleaning for rows to be removed
HRH_clean <- HRH_clean %>%
  filter(!program == "Error")

```

## `r params$fiscal_year` HRH Inventory Completeness and Data Quality
This table shows the number of mechanisms (and their associated staffing expenditure) in `r params$operating_unit` that submitted both the ER and HRH inventory templates this year. This allows us to better understand the overall completeness of the HRH inventory in terms of number of mechanisms that submitted, and the total dollar value that was reported to HRH compared to ER. 

Ideally, the total staffing expenditure reported in HRH and ER by USAID, CDC, and other agencies should be as close to 100% as possible. Due to some inherent reporting differences, the reported HRH and ER staffing expenditures will not always match 100%, but should still be very close. If there are any missing mechanisms that did not submit their HRH inventory template, please follow up with partners accordingly to ensure timely and accurate submission in the next reporting cycle.

```{r, echo = FALSE, eval = TRUE}

# Create summary df to be used for table output
mech_submissions <- finalMerge %>%
  filter(year == 2023) %>%
  filter(GHSC_UN_keywordflags == "FALSE") %>% # excluding all GHSC and UN mechanisms
  filter(operating_unit == params$operating_unit) %>%
  group_by(year, operating_unit, funding_agency, mech_code) %>%
  summarise(HRH_expenditure_amt = sum(HRH_expenditure_amt, na.rm = T), 
            ER_expenditure_amt = sum(ER_expenditure_amt[HRH_relevant == "Y"], na.rm = T)) %>%
  ungroup() %>%
  group_by(funding_agency) %>%
  summarise(Num_HRH_submissions = length(mech_code[HRH_expenditure_amt > 0]),
            Num_ER_submissions = length(mech_code[ER_expenditure_amt > 0]),
            HRH_expenditure_amt = sum(HRH_expenditure_amt, na.rm = T),
            ER_expenditure_amt = sum(ER_expenditure_amt, na.rm = T)) %>%
  ungroup() %>%
  arrange(desc(HRH_expenditure_amt)) %>%
  adorn_totals() %>%
  mutate(pct_submission_rate = HRH_expenditure_amt / ER_expenditure_amt) %>%
  
  # format the columns
  mutate(across(4:5, formattable::currency,digits=0)) %>%
  mutate(across(6, formattable::percent, digits=0))

# Now display the final table with formatted column names
knitr::kable(mech_submissions, "simple", 
             col.names=c("Funding Agency", "No. of HRH Submissions", "No. of ER Submissions", 
                         "HRH Expenditure Amount (USD)", "ER Expenditure Amount   (USD)", "HRH Expenditure as % of ER Expenditure"),
             caption = "Submission Rate by Agency")


```

## `r params$fiscal_year` PEPFAR vs USAID HRH Staffing footprint
This provides a summary of USAID's staffing size relative to the overall PEPFAR program in `r params$operating_unit`. Relative to PEPFAR, this shows USAID's relative staffing footprint (individual count and full time equivalence), the % of spending going towards staffing, and the amount spent by local prime implementation partners. This can serve as a proxy for USAID's role and level of involvement for implementing PEPFAR within the country based on the agency's staffing footprint relative to other agencies such as CDC. 

```{r echo = FALSE, eval = TRUE}

# Create the first df that summarizes PEPFAR
infographic_PEPFAR <- HRH_clean %>%
  filter(fiscal_year == 2023) %>%
  filter(operating_unit == params$operating_unit) %>%
  mutate(funding_agency = "PEPFAR") %>%
  group_by(funding_agency) %>%
  summarise(HRH_expenditure = sum(actual_annual_spend, na.rm = T),
            individual_count = sum(individual_count, na.rm = T),
            annual_fte = sum(annual_fte, na.rm = T),
            placeholder_pctHRH = sum(actual_annual_spend, na.rm = T),
            local_partner_expenditures = sum(actual_annual_spend[is_indigenous_prime_partner == "Y"]))

# Calculate the % spent towards staffing
PEPFAR_pct_HRH <- finalMerge %>%
  filter(year == 2023) %>%
  filter(operating_unit == params$operating_unit) %>%
  group_by(year, operating_unit) %>%
  summarise(HRH_expenditure_amt = sum(HRH_expenditure_amt, na.rm = T),
            ER_expenditure_amt = sum(ER_expenditure_amt, na.rm = T)) %>%
  ungroup() %>%
  mutate(PEPFAR_pct_HRH = HRH_expenditure_amt / ER_expenditure_amt) %>%
  pull(PEPFAR_pct_HRH)

# Replace the % HRH placeholder value with the calculated value
infographic_PEPFAR <- infographic_PEPFAR %>%
  mutate(placeholder_pctHRH = PEPFAR_pct_HRH) %>%
  rename(pct_HRH = placeholder_pctHRH)

# Now, create the second df that summarizes USAID only
infographic_USAID <- HRH_clean %>%
  filter(fiscal_year == 2023) %>%
  filter(operating_unit == params$operating_unit) %>%
  filter(funding_agency == "USAID") %>%
  group_by(funding_agency) %>%
  summarise(HRH_expenditure = sum(actual_annual_spend, na.rm = T),
            individual_count = sum(individual_count, na.rm = T),
            annual_fte = sum(annual_fte, na.rm = T),
            placeholder_pctHRH = sum(actual_annual_spend, na.rm = T),
            local_partner_expenditures = sum(actual_annual_spend[is_indigenous_prime_partner == "Y"]))

# Calculate the % spent towards staffing for USAID
USAID_pct_HRH <- finalMerge %>%
  filter(year == 2023) %>%
  filter(operating_unit == params$operating_unit) %>%
  filter(funding_agency == "USAID") %>%
  group_by(year, operating_unit) %>%
  summarise(HRH_expenditure_amt = sum(HRH_expenditure_amt, na.rm = T),
            ER_expenditure_amt = sum(ER_expenditure_amt, na.rm = T)) %>%
  ungroup() %>%
  mutate(USAID_pct_HRH = HRH_expenditure_amt / ER_expenditure_amt) %>%
  pull(USAID_pct_HRH)

# Replace the % HRH placeholder value with the calculated value for USAID
infographic_USAID <- infographic_USAID %>%
  mutate(placeholder_pctHRH = USAID_pct_HRH) %>%
  rename(pct_HRH = placeholder_pctHRH)

# Lastly, bind both data frames together to create the final summary table
infographic <- rbind(infographic_PEPFAR, infographic_USAID) %>%

# format the columns
  mutate(across(2, formattable::currency, digits=0)) %>%
  mutate(across(3:4, formattable::accounting, digits=0)) %>%
  mutate(across(5, formattable::percent, digits=0)) %>%
  mutate(across(6, formattable::currency, digits=0))

# Display the final summary table with formatted column names
knitr::kable(infographic, "simple", 
             col.names=c("Funding Agency", "HRH Staffing Expenditure", "Individual Count", 
                         "Total FTE (Full-Time Equivalence)", "% of Total Expenditures Spent Towards HRH", 
                         "Local Partner HRH Staffing Expenditure"),
             caption = "HRH Summary: By The Numbers")

```

## Funding Agency
Breakdown of HRH expenditure and number of individual staff supported by USAID, CDC, and other agencies. This provides better understanding of how many staff (and how much they cost) that USAID and CDC recruited to implement their `r params$fiscal_year` program activities.  


```{r fig1, fig.height = 5, fig.width = 7, echo = FALSE, eval = TRUE}

# First, create a df that shows HRH expenditure and individual count by program area
patchwork <- HRH_clean %>%
  filter(fiscal_year == 2023) %>%
  filter(operating_unit == params$operating_unit) %>%
  select(fiscal_year, funding_agency, actual_annual_spend, individual_count) %>%
  group_by(funding_agency) %>%
  summarise(actual_annual_spend = sum(actual_annual_spend, na.rm = T),
            individual_count = sum(individual_count, na.rm = T)) %>%
  arrange(desc(individual_count)) 

# Make the first plot
g1 <- ggplot(data = patchwork, mapping = aes(x = reorder(funding_agency, -actual_annual_spend), y = individual_count, fill = funding_agency)) +
    geom_col(alpha = 0.9,
        width = 0.7) +
  scale_fill_manual(values = c("CDC" = "#f2bc40", 
                              "USAID" = "#c43d4d", 
                              "Other agencies" = "#808080")) +
  
    geom_label(
        aes(y     = individual_count,
            label = scales::comma(round(individual_count))),
        size = 3
    ) +
  
    theme_tq() +
  
    labs(title = stringr::str_glue("Individual Count by Funding Agency in {params$operating_unit}")) +
    theme(axis.line=element_blank(),
        axis.text.y=element_blank(),
        axis.ticks=element_blank(),
        axis.title.x=element_blank(),
        axis.title.y=element_blank(),
        legend.position    = "none",
        plot.title = element_text(size = 10)) 
  
# Make the second plot 

g2 <- ggplot(data = patchwork, mapping = aes(x = reorder(funding_agency, -actual_annual_spend), y = actual_annual_spend, fill = funding_agency)) +
    geom_col(alpha = 0.9,
        width = 0.7) +
  scale_fill_manual(values = c("CDC" = "#f2bc40", 
                              "USAID" = "#c43d4d", 
                              "Other agencies" = "#808080")) +
  
      geom_label(
        aes(y     = actual_annual_spend,
            label = scales::dollar(round(actual_annual_spend))),
        size = 3
    ) +
  
    theme_tq() +
  
    labs(title = stringr::str_glue("HRH Expenditure by Funding Agency in {params$operating_unit}")) +
    theme(axis.line=element_blank(),
        axis.text.y=element_blank(),
        axis.ticks=element_blank(),
        axis.title.x=element_blank(),
        axis.title.y=element_blank(),
        legend.position    = "none",
        plot.title = element_text(size = 10)) 
   
g3 <- g2 + g1

g3


```

\pagebreak

## Primary Program Areas
Breakdown of total HRH expenditure and number of individual staff by primary program area for USAID. This shows a general view of which program areas that staff are 'primarily' working on. Note that the HRH inventory collects info on 'primary' program area in that we ask staff to be categorized towards one program area only. While staff may be working on multiple program areas, staff can only be categorized towards one program area. 

With this in mind, we typically see higher Program Management (PM) and Above Service Programming (ASP) HRH expenditures while seeing lower HRH expenditures for Prevention (PREV), Socioeconomic (SE), or HIV Testing (HTS). This may indicate that non-service delivery staff may be more expensive (though lower by staff count) compared to service delivery staff working at community level (i.e. high staff count, but less expensive)

```{r fig2, fig.height = 7, fig.width = 9, echo = FALSE, eval = TRUE}

# Create a df that shows HRH expenditure and individual count by program area
patchwork <- HRH_clean %>%
  filter(fiscal_year == 2023) %>%
  filter(funding_agency == "USAID") %>%
  filter(operating_unit == params$operating_unit) %>%
  select(fiscal_year, funding_agency, program, actual_annual_spend, individual_count) %>%
  group_by(program) %>%
  summarise(actual_annual_spend = sum(actual_annual_spend, na.rm = T),
            individual_count = sum(individual_count, na.rm = T)) %>%
  arrange(desc(individual_count)) 

# Create % of total table for inline text later
patchwork_pct_PM <- patchwork %>%
  adorn_percentages("col") %>%
  mutate(actual_annual_spend = round(actual_annual_spend * 100),
         individual_count = round(individual_count * 100)) %>%
  filter(program == "PM") %>%
  pull(actual_annual_spend)

# Make the first plot

g1 <- ggplot(data = patchwork, mapping = aes(x = reorder(program, -actual_annual_spend), y = individual_count, fill = program)) +
    geom_col(alpha = 0.9,
        width = 0.7) +
  
    scale_fill_manual(values = c("PM" = "#c43d4d", 
                              "ASP" = "#e07653", 
                              "C&T" = "#2057a7",
                              "SE" = "#1e87a5",
                              "HTS" = "#8980cb",
                              "PREV" = "#808080")) +
  
    geom_label(
        aes(y     = individual_count,
            label = scales::comma(round(individual_count))),
        size = 3
    ) +
  
    theme_tq() +
  
    labs(title = stringr::str_glue("Individual Count by Program Area in {params$operating_unit}")) +
    theme(axis.line=element_blank(),
        axis.text.y=element_blank(),
        axis.ticks=element_blank(),
        axis.title.x=element_blank(),
        axis.title.y=element_blank(),
        legend.position    = "none",
        plot.title = element_text(size = 10)) 

# Make the second plot

g2 <- ggplot(data = patchwork, mapping = aes(x = reorder(program, -actual_annual_spend), y = actual_annual_spend, fill = program)) +
    geom_col(alpha = 0.9,
        width = 0.7) +
  
    scale_fill_manual(values = c("PM" = "#c43d4d", 
                              "ASP" = "#e07653", 
                              "C&T" = "#2057a7",
                              "SE" = "#1e87a5",
                              "HTS" = "#8980cb",
                              "PREV" = "#808080")) +
  
      geom_label(
        aes(y     = actual_annual_spend,
            label = scales::dollar(round(actual_annual_spend))),
        size = 3
    ) +
  
    theme_tq() +
  
    labs(title = stringr::str_glue("HRH Expenditure by Funding Agency in {params$operating_unit}")) +
    theme(axis.line=element_blank(),
        axis.text.y=element_blank(),
        axis.ticks=element_blank(),
        axis.title.x=element_blank(),
        axis.title.y=element_blank(),
        legend.position    = "none",
        plot.title = element_text(size = 10)) 
   
g3 <- g2 + g1

g3

```

\pagebreak

## Top Employment Titles Based on Individual Count
This shows the top 5 employment titles being recruited by USAID based on individual count. The most common employment titles with the highest staff count are typically community-based staff providing direct service delivery. 

```{r fig3, fig.height = 4, fig.width = 7, echo = FALSE, eval = TRUE}

# Create a df that shows individual count by the top 10 employment title
patchwork <- HRH_clean %>%
  filter(fiscal_year == 2023) %>%
  filter(funding_agency == "USAID") %>%
  filter(operating_unit == params$operating_unit) %>%
  select(fiscal_year, funding_agency, employment_title, actual_annual_spend, individual_count) %>%
  group_by(employment_title) %>%
  summarise(actual_annual_spend = sum(actual_annual_spend, na.rm = T),
            individual_count = sum(individual_count, na.rm = T)) %>%
  arrange(desc(individual_count)) %>%
  top_n(n = 5, wt = individual_count) %>%
  ungroup() %>%
  mutate(employment_title_wrapped = str_wrap(employment_title, width = 10)) %>%
  select(-employment_title)

# Make the first plot
g1 <- ggplot(data = patchwork, mapping = aes(x = reorder(employment_title_wrapped, -actual_annual_spend), y = individual_count, fill = employment_title_wrapped)) +
    geom_col(alpha = 0.9,
        width = 0.7) +
  
    scale_fill_manual(values = c("#c43d4d", 
                              "#e07653", 
                              "#2057a7",
                              "#1e87a5",
                              "#8980cb",
                              "#808080")) +
  
    geom_label(
        aes(y     = individual_count,
            label = scales::comma(round(individual_count))),
        size = 3
    ) +
  
    theme_tq() +
  
    labs(title = stringr::str_glue("Individual Count by Top Employment Titles (based on Individual Count)")) +
    theme(axis.line=element_blank(),
        axis.text.y=element_blank(),
        axis.ticks=element_blank(),
        axis.title.x=element_blank(),
        axis.title.y=element_blank(),
        legend.position    = "none",
        plot.title = element_text(size = 9)) 

# Make the second plot


g2 <- ggplot(data = patchwork, mapping = aes(x = reorder(employment_title_wrapped, -actual_annual_spend), y = actual_annual_spend, fill = employment_title_wrapped)) +
    geom_col(alpha = 0.9,
        width = 0.7) +
  
    scale_fill_manual(values = c("#c43d4d", 
                              "#e07653", 
                              "#2057a7",
                              "#1e87a5",
                              "#8980cb",
                              "#808080")) +
  
      geom_label(
        aes(y     = actual_annual_spend,
            label = scales::dollar(round(actual_annual_spend))),
        size = 3
    ) +
  
    theme_tq() +
  
    labs(title = stringr::str_glue("HRH Expenditure by Top Employment Titles (based on Individual Count)")) +
    theme(axis.line=element_blank(),
        axis.text.y=element_blank(),
        axis.ticks=element_blank(),
        axis.title.x=element_blank(),
        axis.title.y=element_blank(),
        legend.position    = "none",
        plot.title = element_text(size = 9)) 
   
g3 <- g2 + g1

g3

```

## Top Employment Titles Based on HRH Expenditure
This shows the top 5 employment titles being recruited by USAID based on HRH expenditure. The most common employment titles with the highest HRH expenditure are typically above site, non-service delivery staff

```{r fig4, fig.height = 4, fig.width = 7, echo = FALSE, eval = TRUE}

# Create a df that shows HRH expenditure by the top 5 employment title
patchwork <- HRH_clean %>%
  filter(fiscal_year == 2023) %>%
  filter(funding_agency == "USAID") %>%
  filter(operating_unit == params$operating_unit) %>%
  select(fiscal_year, funding_agency, employment_title, actual_annual_spend, individual_count) %>%
  group_by(employment_title) %>%
  summarise(actual_annual_spend = sum(actual_annual_spend, na.rm = T),
            individual_count = sum(individual_count, na.rm = T)) %>%
  arrange(desc(actual_annual_spend)) %>%
  top_n(n = 5, wt = actual_annual_spend) %>% # filter for the top 5
  ungroup() %>%
  mutate(employment_title_wrapped = str_wrap(employment_title, width = 10)) %>%
  select(-employment_title)

# Make the first plot
g1 <- ggplot(data = patchwork, mapping = aes(x = reorder(employment_title_wrapped, -actual_annual_spend), y = individual_count, fill = employment_title_wrapped)) +
    geom_col(alpha = 0.9,
        width = 0.7) +
  
    scale_fill_manual(values = c("#c43d4d", 
                              "#e07653", 
                              "#2057a7",
                              "#1e87a5",
                              "#8980cb",
                              "#808080")) +
  
    geom_label(
        aes(y     = individual_count,
            label = scales::comma(round(individual_count))),
        size = 3
    ) +
  
    theme_tq() +
  
    labs(title = stringr::str_glue("Individual Count by Top Employment Titles (based on HRH Expenditure)")) +
    theme(axis.line=element_blank(),
        axis.text.y=element_blank(),
        axis.ticks=element_blank(),
        axis.title.x=element_blank(),
        axis.title.y=element_blank(),
        legend.position    = "none",
        plot.title = element_text(size = 9)) 

# Make the second plot

g2 <- ggplot(data = patchwork, mapping = aes(x = reorder(employment_title_wrapped, -actual_annual_spend), y = actual_annual_spend, fill = employment_title_wrapped)) +
    geom_col(alpha = 0.9,
        width = 0.7) +
  
    scale_fill_manual(values = c("#c43d4d", 
                              "#e07653", 
                              "#2057a7",
                              "#1e87a5",
                              "#8980cb",
                              "#808080")) +
  
      geom_label(
        aes(y     = actual_annual_spend,
            label = scales::dollar(round(actual_annual_spend))),
        size = 3
    ) +
  
    theme_tq() +
  
    labs(title = stringr::str_glue("HRH Expenditure by Top Employment Titles (based on HRH Expenditure)")) +
    theme(axis.line=element_blank(),
        axis.text.y=element_blank(),
        axis.ticks=element_blank(),
        axis.title.x=element_blank(),
        axis.title.y=element_blank(),
        legend.position    = "none",
        plot.title = element_text(size = 9)) 
   
g3 <- g2 + g1

g3

```

## Median Annual Remuneration
This scatter plot shows the median annual remuneration (i.e. estimated annual salary) of the top 20 employment titles among USAID IP's. Each dot is an individual's estimated annual remuneration. This graph generally shows the general spread or range of remuneration for each employment title. 

This graph typically shows that the highest paid staff are typically above site, non-service delivery staff while the lowest paid staff are typically more community-based service delivery staff.

```{r fig5, fig.height = 11, fig.width = 10, echo = FALSE, eval = TRUE}

# Prepare the remuneration df to be used
remuneration <- HRH_clean %>%
  filter(fiscal_year == 2023) %>%
  filter(funding_agency == "USAID") %>%
  filter(operating_unit == params$operating_unit) %>%
  select(employment_title, equivalent_annual_salary)

# Remove any extreme outliers, since we do not want to visualize them in our box plot
detect_outlier <- function(x) {
  # calculate first quantile
  Quantile1 <- quantile(x, probs=.01)
  # calculate third quantile
  Quantile3 <- quantile(x, probs=.98)
  # calculate inter quartile range
  IQR = Quantile3-Quantile1
  # return true or false
  x > Quantile3 + (IQR*1.5) | x < Quantile1 - (IQR*1.5)
}
 
# create remove outlier function
remove_outlier <- function(dataframe,
                            columns=names(dataframe)) {
  # for loop to traverse in columns vector
  for (col in columns) {
    # remove observation if it satisfies outlier function
    dataframe <- dataframe[!detect_outlier(dataframe[[col]]), ]
  }
  
  return(dataframe)
}

# Now use the function to remove extreme outliers from our remuneration df
remuneration <- remove_outlier(remuneration, 'equivalent_annual_salary')

# Calculate the median annual remuneration by employment title, filter for top 10 titles, pull the titles, and use this as a sorting variable
sort_list <- remuneration %>%
  group_by(employment_title) %>%
  summarise(median_remuneration = median(equivalent_annual_salary, na.rm = T),
            n = n()) %>%
  arrange(desc(median_remuneration)) %>%
  filter(n > 10) %>%
  top_n(n = 20, wt = median_remuneration) %>%
  mutate(employment_title_wrapped = str_wrap(employment_title, width = 10)) %>%
  pull(employment_title)
  
# Filter our remuneration df for the top titles in our sorting list
remuneration <- remuneration %>%
  filter(employment_title %in% sort_list)

# Now create the box plot
ggplot(remuneration, 
       aes(x=reorder(employment_title, equivalent_annual_salary, FUN = median), 
           y=equivalent_annual_salary)) + 
  
  geom_jitter(aes(colour = employment_title), width = 0.3, shape = 21, color = "#002a6c") +
  
  stat_summary(fun = median, show.legend = FALSE, geom = "crossbar", color = "black", width = 0.8) +
  
  scale_y_continuous(labels = label_dollar()) +
  
  coord_flip() + 
  
  labs(x = "Employment Title", y = "Median Annual Remuneration") + 
  
  scale_fill_manual(values = rep(NA, 2)) +
  
  theme_tq() +
  
  theme(panel.background = element_rect(fill = "white"),
        axis.line=element_blank(),
        axis.ticks=element_blank(),
        axis.title.y=element_blank(),
        plot.title = element_text(size = 10),
        legend.position    = "none")

```

## Local Partners
This shows how much of USAID's HRH expenditure are channeled through prime local partners for each primary program area. Note that Local Partners here are defined as local organizations that were awarded as the prime partner, and not as sub-awardees/subrecipients. If a local organization is a subrecipient and the prime partner is an international organization, then their staffing expenditures are considered as non-local

This should provide country teams a better idea of the level of local partner involvement in implementing PEPFAR program activities, especially as prime partners

```{r fig6, fig.height = 5, fig.width = 6, echo = FALSE, eval = TRUE}

# Create the summary df for local partners
local_partner <- HRH_clean %>%
  filter(fiscal_year == 2023) %>%
  filter(funding_agency == "USAID") %>%
  filter(operating_unit == params$operating_unit) %>%
  select(fiscal_year, funding_agency, program, is_indigenous_prime_partner, actual_annual_spend) %>%
  group_by(program, is_indigenous_prime_partner) %>%
  summarise(actual_annual_spend = sum(actual_annual_spend, na.rm = T)) %>%
  arrange(desc(actual_annual_spend)) 

# Now plot the summary df
  ggplot(local_partner, aes(x = reorder(program, -actual_annual_spend), 
                          y = actual_annual_spend, 
                          fill = is_indigenous_prime_partner,
                          label = actual_annual_spend)) +
  
  geom_bar(stat = "identity") +
  
  geom_text(aes(label = scales::dollar(round(actual_annual_spend))), size = 3, position = position_stack(vjust = 0.5)) +
  
  xlab("Program Area") +
    
  ylab("Total HRH Expenditure") +
    
  labs(fill = "Prime Local Partner?") +
    
  scale_y_continuous(labels = label_dollar()) +
    
  scale_fill_manual(values = c("Y" = "#1e87a5",
                              "N" = "#808080")) +

  theme(panel.background = element_rect(fill = "white"),
        axis.line=element_blank(),
        axis.text.y=element_blank(),
        axis.ticks=element_blank(),
        axis.title.x=element_blank(),
        axis.title.y=element_blank(),
        plot.title = element_text(size = 10)) +
  
  theme_tq() 
  
```

\pagebreak

## Work Location
This shows the work location of USAID staff in terms of HRH expenditure and staff count. Most staff are typically located at the community level and facility level. There are typically less staff working above site (i.e. non-service delivery staff), but typically associated with higher HRH expenditure. 

Above site staff are defined as personnel that are not directly interacting or providing services to beneficiaries (e.g. TA personnel or support staff based mostly in offices). Roving staff are direct service delivery staff that are providing services in multiple facilities.

```{r fig7, fig.height = 7, fig.width = 9, echo = FALSE, eval = TRUE}

# Create a df that shows HRH expenditure and individual count by work location
patchwork <- HRH_clean %>%
  filter(fiscal_year == 2023) %>%
  filter(funding_agency == "USAID") %>%
  filter(operating_unit == params$operating_unit) %>%
  select(fiscal_year, work_location, actual_annual_spend, individual_count) %>%
  group_by(work_location) %>%
  summarise(actual_annual_spend = sum(actual_annual_spend, na.rm = T),
            individual_count = sum(individual_count, na.rm = T)) %>%
  arrange(desc(individual_count)) 

# Make the Y-Axis Transformer to properly scale / align the dual axis in the plot
transformer <- patchwork %>%
    transformer_dual_y_axis(
        primary_column   = individual_count,
        secondary_column = actual_annual_spend,
        include_y_zero   = TRUE
    )

# Now make the plot
g1 <- patchwork %>%
    ggplot(aes(x = reorder(work_location, -individual_count))) +
    geom_col(
        aes(y = individual_count, fill = "Individual Count"),
        alpha = 0.9,
        width = 0.7
    ) +
  
    geom_label(
        aes(
            y     = individual_count,
            label = scales::comma(round(individual_count)),
            color = "Individual Count",
        ),
        size = 4
    )

# ADD SECONDARY AXIS ----

# Create a y-label distance for use in this graph based on the max value of the y-axis
y_axis_nudge <- patchwork %>%
  summarise(max_y_value = max(actual_annual_spend)) %>%
  mutate(nudge = max_y_value * .000025) %>%
  pull(nudge)

g2 <- g1 +
    geom_point(
        aes(y     = transformer$inv_func(actual_annual_spend),
            color = "HRH Expenditure"),
        size = 4
    ) +
  
    geom_label(
        aes(
            y     = transformer$inv_func(actual_annual_spend),
            label = scales::dollar(round(actual_annual_spend)),
            color = "HRH Expenditure"),
        size = 3,
        nudge_y = y_axis_nudge
    ) + 
  
    scale_y_continuous(
        name     = "Individual Count",
        labels = scales::comma,
        sec.axis = sec_axis(
            trans = ~ transformer$scale_func(.),
            name  = "Total HRH Expenditure",
            labels = scales::dollar
        )) +
  
  labs(title = stringr::str_glue("Individual Count and HRH Expenditure by Work Location in {params$operating_unit}")) +
  
  theme(panel.grid = element_blank(), panel.background = element_rect(fill = "white"),
        axis.line=element_blank(),
        axis.text.y=element_blank(),
        axis.ticks=element_blank(),
        axis.title.x=element_blank(),
        axis.title.y=element_blank(),
        plot.title = element_text(size = 10)) 

g3 <- g2 +
    theme_tq() +
    scale_color_manual(values = c(
        palette_light()[["red"]],
        palette_light()[["blue"]]
    )) +
    scale_fill_manual(values = palette_light()[["blue"]]) +
    theme(
        legend.position    = "none",
        axis.title.y.right = element_text(color = palette_light()[["red"]]),
        axis.text.y.right  = element_text(color = palette_light()[["red"]]),
        axis.title.x=element_blank()
    )

g3

```

\pagebreak

## Top Employment Titles for Community-Based Staff
This shows the most common employment title for community-based staff by count

```{r fig8, fig.height = 5, fig.width = 8, echo = FALSE, eval = TRUE}

# Create a df that shows HRH expenditure and individual count by the top 5 community titles
patchwork <- HRH_clean %>%
  filter(fiscal_year == 2023) %>%
  filter(funding_agency == "USAID") %>%
  filter(operating_unit == params$operating_unit) %>%
  filter(is_community_primarily == "Yes") %>%
  select(fiscal_year, work_location, employment_title, actual_annual_spend, individual_count) %>%
  group_by(employment_title) %>%
  summarise(actual_annual_spend = sum(actual_annual_spend, na.rm = T),
            individual_count = sum(individual_count, na.rm = T)) %>%
  top_n(n = 5, wt = individual_count) %>%
  ungroup() %>%
  mutate(employment_title_wrapped = str_wrap(employment_title, width = 10)) %>%
  select(-employment_title)

# Make the column graph
g1 <- patchwork %>%
    ggplot(aes(x = reorder(employment_title_wrapped, individual_count))) +
    geom_col(
        aes(y = individual_count, fill = "Individual Count"),
        alpha = 0.9,
        width = 0.8
    ) +
    geom_label(
        aes(
            y     = individual_count,
            label = scales::comma(round(individual_count)),
            color = "Individual Count",
        ),
        size = 3
    ) +
  
  ylab("Individual Count") + 
  
  theme(panel.grid = element_blank(), panel.background = element_rect(fill = "white"),
        axis.line=element_blank(),
        axis.text.y=element_blank(),
        axis.ticks=element_blank(),
        axis.title.x=element_blank(),
        axis.title.y=element_blank(),
        plot.title = element_text(size = 10)) +
  
    theme_tq() +
  
    scale_color_manual(values = c(
        palette_light()[["blue"]]
    )) +
  
    scale_fill_manual(values = palette_light()[["blue"]]) +
    theme(
        legend.position    = "none",
        axis.title.y=element_blank()
    ) +

    scale_y_continuous(labels = scales::comma_format()) +
  
    coord_flip()

g1

```

## Top Primary Beneficiaries for Community-Based Staff
This shows the top primary beneficiaries supported by community-based staff by count

```{r fig9, fig.height = 5, fig.width = 9, echo = FALSE, eval = TRUE}

# Create a df that shows HRH expenditure and individual count by primary beneficiary
patchwork <- HRH_clean %>%
  filter(fiscal_year == 2023) %>%
  filter(funding_agency == "USAID") %>%
  filter(operating_unit == params$operating_unit) %>%
  filter(work_location == "Community") %>%
  select(fiscal_year, beneficiary, actual_annual_spend, individual_count) %>%
  mutate(beneficiary_wrapped = str_wrap(beneficiary, width = 10)) %>%
  group_by(beneficiary_wrapped) %>%
  summarise(actual_annual_spend = sum(actual_annual_spend, na.rm = T),
            individual_count = sum(individual_count, na.rm = T)) %>%
  arrange(desc(individual_count)) 

# Make A Y-Axis Transformer
transformer <- patchwork %>%
    transformer_dual_y_axis(
        primary_column   = individual_count,
        secondary_column = actual_annual_spend,
        include_y_zero   = TRUE
    )

# Make the first plot

g1 <- patchwork %>%
    ggplot(aes(x = reorder(beneficiary_wrapped, -individual_count))) +
    geom_col(
        aes(y = individual_count, fill = "Individual Count"),
        alpha = 0.9,
        width = 0.8
    ) +
    geom_label(
        aes(
            y     = individual_count,
            label = scales::comma(round(individual_count)),
            color = "Individual Count",
        ),
        size = 3
    )

# ADD SECONDARY AXIS ----

# Create a y-label distance for use in this graph based on the max value of the y-axis
y_axis_nudge <- patchwork %>%
  summarise(max_y_value = max(actual_annual_spend)) %>%
  mutate(nudge = max_y_value * .00015) %>%
  pull(nudge)

g2 <- g1 +
    geom_point(
        aes(y     = transformer$inv_func(actual_annual_spend),
            color = "HRH Expenditure"),
        size = 4
    ) +
    geom_label(
        aes(
            y     = transformer$inv_func(actual_annual_spend),
            label = scales::dollar(round(actual_annual_spend)),
            color = "HRH Expenditure"),
        size = 3,
        nudge_y = y_axis_nudge
    ) + 
  

    scale_y_continuous(
        name     = "Individual Count",
        labels = scales::comma,
        sec.axis = sec_axis(
            trans = ~ transformer$scale_func(.),
            name  = "Total HRH Expenditure",
            labels = scales::dollar
        )
    ) +
  
  theme(panel.grid = element_blank(), panel.background = element_rect(fill = "white"),
        axis.line=element_blank(),
        axis.text.y=element_blank(),
        axis.ticks=element_blank(),
        axis.title.x=element_blank(),
        axis.title.y=element_blank(),
        plot.title = element_text(size = 10)) 

g3 <- g2 +
    theme_tq() +
    scale_color_manual(values = c(
        palette_light()[["red"]],
        palette_light()[["blue"]]
    )) +
    scale_fill_manual(values = palette_light()[["blue"]]) +
    theme(
        legend.position    = "none",
        axis.title.y.right = element_text(color = palette_light()[["red"]]),
        axis.text.y.right  = element_text(color = palette_light()[["red"]]),
        axis.title.x=element_blank()
    ) 

g3

```

## Top Prime Implementation Partners 
This shows the top 5 prime partners with the highest staffing footprint in terms of individual count. Each prime partner is disaggregated by primary program area, which should indicate the primary program activities that each partner is focused on. 

If requested, we can provide a deeper dive review of each top implementation partner to better understand their staffing composition.

```{r fig10, fig.height = 8, fig.width = 9, echo = FALSE, eval = TRUE}

# Pull names of top 5 partners only
top_5_partners <- HRH_clean %>%
  filter(fiscal_year == 2023) %>%
  filter(funding_agency == "USAID") %>%
  filter(operating_unit == params$operating_unit) %>%
  group_by(prime_partner_name) %>%
  summarise(individual_count = sum(individual_count, na.rm = T)) %>%
  arrange(desc(individual_count)) %>%
  top_n(n = 5, wt = individual_count) %>%
  pull(prime_partner_name)

# Create a summary table that filters for the top 5 partner names only
prime_partners <- HRH_clean %>%
  filter(fiscal_year == 2023) %>%
  filter(funding_agency == "USAID") %>%
  filter(operating_unit == params$operating_unit) %>%
  filter(prime_partner_name %in% top_5_partners) %>%
  select(fiscal_year, er_category, prime_partner_name, individual_count) %>%
  group_by(prime_partner_name, er_category) %>%
  summarise(individual_count = sum(individual_count, na.rm = T)) %>%
  arrange(desc(individual_count)) %>%
  mutate(prime_partner_name_wrapped = str_wrap(prime_partner_name, width = 10))
  
# Now make the plot
  ggplot(prime_partners, aes(x = reorder(prime_partner_name_wrapped, -individual_count), 
                             y = individual_count, 
                             fill = er_category,
                             label = individual_count)) +
  
  geom_bar(stat = "identity") +
  
  geom_text(aes(label = scales::comma(round(individual_count))), 
            size = 4, 
            position = position_stack(vjust = 0.5)) +
    
  scale_y_continuous(labels = scales::comma_format()) +
    
  xlab("Prime partner name") +
    
  ylab("Individual Count") + 
  
  scale_fill_manual(values = c("#287c6f", "#2057a7", "#808080", "#C43d4d")) +
  
  theme(panel.grid = element_blank(), panel.background = element_rect(fill = "white"),
        axis.line=element_blank(),
        axis.text.y=element_blank(),
        axis.ticks=element_blank(),
        axis.title.x=element_blank(),
        axis.title.y=element_blank(),
        plot.title = element_text(size = 10)) +
    
  theme_tq() 

```

