---
output:
  word_document:
    reference_docx: word_template.docx
author: "GH/OHA/SPS/HWF"
geometry: margin = 2cm
date: "`r format(Sys.time(), '%d %B, %Y')`"

params: 
    set_title: "My Title!"
    operating_unit: !r c("Lesotho")
    fiscal_year: "FY24"
---

---
title: "`r params$fiscal_year` Human Resources for Health (HRH) Inventory Summary - `r params$operating_unit`"
---


```{r setup, include=FALSE}

# set global standards
knitr::opts_chunk$set(echo = FALSE) # supress code outputs
knitr::opts_chunk$set(warning = FALSE, message = FALSE) # suppress all warnings
knitr::opts_chunk$set(dpi=300) # set high resolution of plots

# load all needed libraries
library(plyr)
library(dplyr)
library(tidyverse)
library(sqldf)
library(janitor)
library(readxl)
library(writexl)
library(shiny)
library(scales)
library(tidyquant)
library(formattable)
library(kableExtra)
library(ggrepel)
library(ggplot2)
library(patchwork)


```

## Background

This HRH Inventory Briefer provides a detailed summary of the FY24 HRH Inventory for Tanzania. This report will summarize the HRH inventory templates submitted by prime implementation partners to help Mission teams better monitor and understand their PEPFAR-supported staffing footprint.

Missions should use these briefers to review their staffing investments and trends at the OU level to inform programming and facilitate HRH planning. If there are questions on the summary info provided here, please reach out to the HRH Reporting Helpdesk (hrh-reporting-helpdesk@usaid.gov)

```{r}

# Load the FSD, HRH, and merged FSD-HRH dataset
fin_data_orig <- read.delim("./1. Data/Financial_Structured_Datasets_COP17-24_20241213.txt") # read in FSD dataset
load(file = "./4. Outputs/RDS/FY24_cleanHRH.rds") # cleaned HRH dataset
load(file = "./4. Outputs/RDS/HRH_ER_merged_21_24.rds") # HRH-ER merged dataset


```

## `r params$fiscal_year` HRH Inventory Completeness and Data Quality
This table shows the number of mechanisms (and their associated staffing expenditure) in `r params$operating_unit` that submitted both the ER and HRH inventory templates for `r params$fiscal_year`. This allows us to better understand the overall completeness of the HRH inventory in terms of the number of mechanisms that were submitted, and the total dollar value that was reported to HRH compared to ER. 

The total staffing expenditure reported in HRH and ER by USAID, CDC, and other agencies should closely match each other. The reported HRH and ER staffing expenditures will not always exactly match each other, but are still expected to be aligned. HRH and ER reporting from each implementing partner must collaborate together to ensure alignment of reported staffing expenditure. If there are mechanisms that submitted staffing expenses in their Expenditure Report, but did not submit their HRH inventory template, please follow up with partners accordingly to ensure timely submissions in the next reporting cycle.

```{r, echo = FALSE, eval = TRUE}

# Create summary df to be used for table output
mech_submissions <- finalMerge %>%
  filter(year == 2024) %>%
  filter(UN_keywordflags == "FALSE") %>% # excluding all GHSC and UN mechanisms
  filter(operating_unit == params$operating_unit) %>%
  group_by(year, operating_unit, funding_agency, mech_code) %>%
  summarise(HRH_expenditure_amt = sum(HRH_expenditure_amt, na.rm = T), 
            ER_expenditure_amt = sum(ER_expenditure_amt[HRH_relevant == "Y"], na.rm = T)) %>%
  ungroup() %>%
  group_by(funding_agency) %>%
  summarise(Num_HRH_submissions = length(mech_code[HRH_expenditure_amt > 0]),
            Num_ER_submissions = length(mech_code[ER_expenditure_amt > 0]),
            HRH_expenditure_amt = sum(HRH_expenditure_amt, na.rm = T),
            ER_expenditure_amt = sum(ER_expenditure_amt, na.rm = T)) %>%
  ungroup() %>%
  arrange(desc(HRH_expenditure_amt)) %>%
  adorn_totals() %>%
  mutate(pct_submission_rate = HRH_expenditure_amt / ER_expenditure_amt) %>%
  
  # format the columns
  mutate(across(4:5, formattable::currency,digits=0)) %>%
  mutate(across(6, formattable::percent, digits=0))

# Now display the final table with formatted column names
knitr::kable(mech_submissions, "simple", 
             col.names=c("Funding Agency", "No. of HRH Submissions", "No. of ER Submissions", 
                         "HRH Expenditure Amount (USD)", "ER Expenditure Amount   (USD)", "HRH Expenditure as % of ER Expenditure"),
             caption = "Submission Rate by Agency")

#mech_submissions <- gt(mech_submissions)


```

## `r params$fiscal_year` PEPFAR vs USAID HRH Staffing footprint

This provides a summary of USAID’s staffing size relative to the overall PEPFAR program in `r params$operating_unit`. This shows USAID’s staffing footprint relative to the full PEPFAR staffing footprint (individual count and full time equivalence), the % of spending going towards staffing, and the amount spent by local prime implementation partners. This serves as an indicator for USAID’s level of staffing investment in the country relative to other implementing agencies.

```{r echo = FALSE, eval = TRUE}

# Create the first df that summarizes PEPFAR
infographic_PEPFAR <- HRH_clean %>%
  filter(fiscal_year == 2024) %>%
  filter(operating_unit == params$operating_unit) %>%
  mutate(funding_agency = "PEPFAR") %>%
  group_by(funding_agency) %>%
  summarise(HRH_expenditure = sum(actual_annual_spend, na.rm = T),
            individual_count = sum(individual_count, na.rm = T),
            annual_fte = sum(annual_fte, na.rm = T),
            placeholder_pctHRH = sum(actual_annual_spend, na.rm = T),
            local_partner_expenditures = sum(actual_annual_spend[is_indigenous_prime_partner == "Y"]))

# Calculate the % spent towards staffing
PEPFAR_pct_HRH <- finalMerge %>%
  filter(year == 2024) %>%
  filter(operating_unit == params$operating_unit) %>%
  group_by(year, operating_unit) %>%
  summarise(HRH_expenditure_amt = sum(HRH_expenditure_amt, na.rm = T),
            ER_expenditure_amt = sum(ER_expenditure_amt, na.rm = T)) %>%
  ungroup() %>%
  mutate(PEPFAR_pct_HRH = HRH_expenditure_amt / ER_expenditure_amt) %>%
  pull(PEPFAR_pct_HRH)

# Replace the % HRH placeholder value with the calculated value
infographic_PEPFAR <- infographic_PEPFAR %>%
  mutate(placeholder_pctHRH = PEPFAR_pct_HRH) %>%
  rename(pct_HRH = placeholder_pctHRH)

# Now, create the second df that summarizes USAID only
infographic_USAID <- HRH_clean %>%
  filter(fiscal_year == 2024) %>%
  filter(operating_unit == params$operating_unit) %>%
  filter(funding_agency == "USAID") %>%
  group_by(funding_agency) %>%
  summarise(HRH_expenditure = sum(actual_annual_spend, na.rm = T),
            individual_count = sum(individual_count, na.rm = T),
            annual_fte = sum(annual_fte, na.rm = T),
            placeholder_pctHRH = sum(actual_annual_spend, na.rm = T),
            local_partner_expenditures = sum(actual_annual_spend[is_indigenous_prime_partner == "Y"]))

# Calculate the % spent towards staffing for USAID
USAID_pct_HRH <- finalMerge %>%
  filter(year == 2024) %>%
  filter(operating_unit == params$operating_unit) %>%
  filter(funding_agency == "USAID") %>%
  group_by(year, operating_unit) %>%
  summarise(HRH_expenditure_amt = sum(HRH_expenditure_amt, na.rm = T),
            ER_expenditure_amt = sum(ER_expenditure_amt, na.rm = T)) %>%
  ungroup() %>%
  mutate(USAID_pct_HRH = HRH_expenditure_amt / ER_expenditure_amt) %>%
  pull(USAID_pct_HRH)

# Replace the % HRH placeholder value with the calculated value for USAID
infographic_USAID <- infographic_USAID %>%
  mutate(placeholder_pctHRH = USAID_pct_HRH) %>%
  rename(pct_HRH = placeholder_pctHRH)

# Lastly, bind both data frames together to create the final summary table
infographic <- rbind(infographic_PEPFAR, infographic_USAID) %>%

# format the columns
  mutate(across(2, formattable::currency, digits=0)) %>%
  mutate(across(3:4, formattable::accounting, digits=0)) %>%
  mutate(across(5, formattable::percent, digits=0)) %>%
  mutate(across(6, formattable::currency, digits=0))

# Display the final summary table with formatted column names
knitr::kable(infographic, "simple", 
             col.names=c("Funding Agency", "HRH Staffing Expenditure", "Individual Count", 
                         "Total FTE (Full-Time Equivalence)", "% of Total Expenditures Spent Towards HRH", 
                         "Local Partner HRH Staffing Expenditure"),
             caption = "HRH Summary: By The Numbers")

```

## Funding Agency

Breakdown of HRH expenditure and number of individual staff supported by USAID, CDC, and other agencies. This provides a summary of the total staffing expenditure and total headcount that are being supported by each agency to implement their `r params$fiscal_year` program activities.

```{r fig1, fig.height = 6, fig.width = 10, echo = FALSE, eval = TRUE}

# First, create a df that shows HRH expenditure and individual count by program area
patchwork <- HRH_clean %>%
  filter(fiscal_year == 2024) %>%
  filter(operating_unit == params$operating_unit) %>%
  select(fiscal_year, funding_agency, actual_annual_spend, individual_count) %>%
  group_by(funding_agency) %>%
  summarise(actual_annual_spend = sum(actual_annual_spend, na.rm = T),
            individual_count = sum(individual_count, na.rm = T)) %>%
  arrange(desc(individual_count)) 

# Make the first plot
g1 <- ggplot(data = patchwork, mapping = aes(x = reorder(funding_agency, -actual_annual_spend), y = individual_count, fill = funding_agency)) +
    geom_col(alpha = 0.9,
        width = 0.8) +
  scale_fill_manual(values = c("CDC" = "#f2bc40", 
                              "USAID" = "#c43d4d", 
                              "Other agencies" = "#808080")) +
  
    geom_label(
        aes(y     = individual_count,
            label = scales::comma(round(individual_count))),
        size = 5
    ) +
  
    theme_tq() +
  
    labs(title = stringr::str_glue("Individual Count by Funding Agency")) +
    theme(axis.line=element_blank(),
        axis.text.y=element_blank(),
        axis.ticks=element_blank(),
        axis.title.x=element_blank(),
        axis.title.y=element_blank(),
        legend.position    = "none",
        plot.title = element_text(size = 15),
        text = element_text(size = 18)) 
  
# Make the second plot 

g2 <- ggplot(data = patchwork, mapping = aes(x = reorder(funding_agency, -actual_annual_spend), y = actual_annual_spend, fill = funding_agency)) +
    geom_col(alpha = 0.9,
        width = 0.8) +
  scale_fill_manual(values = c("CDC" = "#f2bc40", 
                              "USAID" = "#c43d4d", 
                              "Other agencies" = "#808080")) +
  
      geom_label(
        aes(y     = actual_annual_spend,
            label = scales::dollar(round(actual_annual_spend))),
        size = 5
    ) +
  
    theme_tq() +
  
    labs(title = stringr::str_glue("HRH Expenditure by Funding Agency")) +
    theme(axis.line=element_blank(),
        axis.text.y=element_blank(),
        axis.ticks=element_blank(),
        axis.title.x=element_blank(),
        axis.title.y=element_blank(),
        legend.position    = "none",
        plot.title = element_text(size = 15),
        text = element_text(size = 18)) 
   
g3 <- g2 + g1

g3


```

\pagebreak

## Primary Program Areas
Breakdown of total HRH expenditure and number of individual staff by primary program area for USAID. This shows a general view of which program areas that staff are ‘primarily’ working on. Note that the HRH inventory collects data on ‘primary’ program areas where staff can only be categorized towards one program area that accounts for most of their worker time. While staff may be working on multiple program areas, staff can only be categorized towards one program area in the HRH inventory.

With this in mind, we typically see higher Program Management (PM) and Above Service Programming (ASP) staffing expenditures while seeing lower staffing expenditures for Prevention (PREV), Socioeconomic (SE), or HIV Testing (HTS). This may indicate that non-service delivery staff may be higher cost (though lower by staff count) compared to service delivery staff working at community level (with higher staff count, but lower cost)

```{r fig2, fig.height = 8, fig.width = 10, echo = FALSE, eval = TRUE}

# Create a df that shows HRH expenditure and individual count by program area
patchwork <- HRH_clean %>%
  filter(fiscal_year == 2024) %>%
  filter(funding_agency == "USAID") %>%
  filter(operating_unit == params$operating_unit) %>%
  select(fiscal_year, funding_agency, program, actual_annual_spend, individual_count) %>%
  group_by(program) %>%
  summarise(actual_annual_spend = sum(actual_annual_spend, na.rm = T),
            individual_count = sum(individual_count, na.rm = T)) %>%
  arrange(desc(individual_count)) 

# Create % of total table for inline text later
patchwork_pct_PM <- patchwork %>%
  adorn_percentages("col") %>%
  mutate(actual_annual_spend = round(actual_annual_spend * 100),
         individual_count = round(individual_count * 100)) %>%
  filter(program == "PM") %>%
  pull(actual_annual_spend)

# Make the first plot

g1 <- ggplot(data = patchwork, mapping = aes(x = reorder(program, -actual_annual_spend), y = individual_count, fill = program)) +
    geom_col(alpha = 0.9,
        width = 0.8) +
  
    scale_fill_manual(values = c("PM" = "#c43d4d", 
                              "ASP" = "#e07653", 
                              "C&T" = "#2057a7",
                              "SE" = "#1e87a5",
                              "HTS" = "#8980cb",
                              "PREV" = "#808080")) +
  
    geom_label_repel(data = patchwork, aes(label = scales::comma(round(individual_count))),
             size = 5,
             position = "identity",
             min.segment.length = 0.25,
             direction = "y",
             show.legend = FALSE) +
  
    theme_tq() +
  
    labs(title = stringr::str_glue("Individual Count by Program Area in {params$operating_unit}")) +
    theme(axis.line=element_blank(),
        axis.text.y=element_blank(),
        axis.ticks=element_blank(),
        axis.title.x=element_blank(),
        axis.title.y=element_blank(),
        legend.position    = "none",
        plot.title = element_text(size = 15),
        text = element_text(size = 18)) 

# Make the second plot

g2 <- ggplot(data = patchwork, mapping = aes(x = reorder(program, -actual_annual_spend), y = actual_annual_spend, fill = program)) +
    geom_col(alpha = 0.9,
        width = 0.8) +
  
    scale_fill_manual(values = c("PM" = "#c43d4d", 
                              "ASP" = "#e07653", 
                              "C&T" = "#2057a7",
                              "SE" = "#1e87a5",
                              "HTS" = "#8980cb",
                              "PREV" = "#808080")) +
  
    geom_label_repel(data = patchwork, aes(label = scales::dollar(round(actual_annual_spend))),
             size = 5,
             position = "identity",
             min.segment.length = 0.25,
             direction = "y",
             show.legend = FALSE) +
  
    theme_tq() +
  
    labs(title = stringr::str_glue("HRH Expenditure by Funding Agency in {params$operating_unit}")) +
    theme(axis.line=element_blank(),
        axis.text.y=element_blank(),
        axis.ticks=element_blank(),
        axis.title.x=element_blank(),
        axis.title.y=element_blank(),
        legend.position    = "none",
        plot.title = element_text(size = 15),
        text = element_text(size = 18)) 
   
g3 <- g2 + g1

g3

```

\pagebreak

## Top Employment Titles Based on Individual Count
This shows the top 5 employment titles supported by PEPFAR USAID based on individual count. The most common employment titles with the highest staff count are typically community-based staff providing direct service delivery.

```{r fig3, fig.height = 4, fig.width = 7, echo = FALSE, eval = TRUE}

# Create a df that shows individual count by the top 10 employment title
patchwork <- HRH_clean %>%
  filter(fiscal_year == 2024) %>%
  filter(funding_agency == "USAID") %>%
  filter(operating_unit == params$operating_unit) %>%
  select(fiscal_year, funding_agency, employment_title, actual_annual_spend, individual_count) %>%
  group_by(employment_title) %>%
  summarise(actual_annual_spend = sum(actual_annual_spend, na.rm = T),
            individual_count = sum(individual_count, na.rm = T)) %>%
  arrange(desc(individual_count)) %>%
  top_n(n = 5, wt = individual_count) %>%
  ungroup() %>%
  mutate(employment_title_wrapped = str_wrap(employment_title, width = 10)) %>%
  select(-employment_title)

# Make the first plot
g1 <- ggplot(data = patchwork, mapping = aes(x = reorder(employment_title_wrapped, -actual_annual_spend), y = individual_count, fill = employment_title_wrapped)) +
    geom_col(alpha = 0.9,
        width = 0.8) +
  
    scale_fill_manual(values = c("#c43d4d", 
                              "#e07653", 
                              "#2057a7",
                              "#1e87a5",
                              "#8980cb",
                              "#808080")) +
  
    geom_label_repel(data = patchwork, aes(label = scales::comma(round(individual_count))),
             size = 3,
             position = "identity",
             min.segment.length = 0.25,
             direction = "y",
             show.legend = FALSE) +
  
    theme_tq() +
  
    labs(title = stringr::str_glue("Individual Count by Top Employment Titles")) +
    theme(axis.line=element_blank(),
        axis.text.y=element_blank(),
        axis.ticks=element_blank(),
        axis.title.x=element_blank(),
        axis.title.y=element_blank(),
        legend.position    = "none",
        plot.title = element_text(size = 10),
        axis.text.x = element_text(angle = 45, vjust = 1, hjust = 1)) 

# Make the second plot

g2 <- ggplot(data = patchwork, mapping = aes(x = reorder(employment_title_wrapped, -actual_annual_spend), y = actual_annual_spend, fill = employment_title_wrapped)) +
    geom_col(alpha = 0.9,
        width = 0.8) +
  
    scale_fill_manual(values = c("#c43d4d", 
                              "#e07653", 
                              "#2057a7",
                              "#1e87a5",
                              "#8980cb",
                              "#808080")) +
  
    geom_label_repel(data = patchwork, aes(label = scales::dollar(round(actual_annual_spend))),
             size = 3,
             position = "identity",
             min.segment.length = 0.25,
             direction = "y",
             show.legend = FALSE) +
  
    theme_tq() +
  
    labs(title = stringr::str_glue("HRH Expenditure by Top Employment Titles")) +
    theme(axis.line=element_blank(),
        axis.text.y=element_blank(),
        axis.ticks=element_blank(),
        axis.title.x=element_blank(),
        axis.title.y=element_blank(),
        legend.position    = "none",
        plot.title = element_text(size = 10),
        axis.text.x = element_text(angle = 45, vjust = 1, hjust = 1)) 
   
g3 <- g2 + g1

g3

```

## Top Employment Titles Based on HRH Expenditure
This shows the top 5 employment titles supported by PEPFAR USAID based on HRH expenditure. The most common employment titles with the highest HRH expenditure are typically above site, non-service delivery staff

```{r fig4, fig.height = 4, fig.width = 7, echo = FALSE, eval = TRUE}

# Create a df that shows HRH expenditure by the top 5 employment title
patchwork <- HRH_clean %>%
  filter(fiscal_year == 2024) %>%
  filter(funding_agency == "USAID") %>%
  filter(operating_unit == params$operating_unit) %>%
  select(fiscal_year, funding_agency, employment_title, actual_annual_spend, individual_count) %>%
  group_by(employment_title) %>%
  summarise(actual_annual_spend = sum(actual_annual_spend, na.rm = T),
            individual_count = sum(individual_count, na.rm = T)) %>%
  arrange(desc(actual_annual_spend)) %>%
  top_n(n = 5, wt = actual_annual_spend) %>% # filter for the top 5
  ungroup() %>%
  mutate(employment_title_wrapped = str_wrap(employment_title, width = 10)) %>%
  select(-employment_title)

# Make the first plot
g1 <- ggplot(data = patchwork, mapping = aes(x = reorder(employment_title_wrapped, -actual_annual_spend), y = individual_count, fill = employment_title_wrapped)) +
    geom_col(alpha = 0.9,
        width = 0.8) +
  
    scale_fill_manual(values = c("#c43d4d", 
                              "#e07653", 
                              "#2057a7",
                              "#1e87a5",
                              "#8980cb",
                              "#808080")) +
  
    geom_label_repel(data = patchwork, aes(label = scales::comma(round(individual_count))),
             size = 3,
             position = "identity",
             min.segment.length = 0.25,
             direction = "y",
             show.legend = FALSE) +
  
    theme_tq() +
  
    labs(title = stringr::str_glue("Individual Count by Top Employment Titles")) +
    theme(axis.line=element_blank(),
        axis.text.y=element_blank(),
        axis.ticks=element_blank(),
        axis.title.x=element_blank(),
        axis.title.y=element_blank(),
        legend.position    = "none",
        plot.title = element_text(size = 10),
        axis.text.x = element_text(angle = 45, vjust = 1, hjust = 1)) 

# Make the second plot

g2 <- ggplot(data = patchwork, mapping = aes(x = reorder(employment_title_wrapped, -actual_annual_spend), y = actual_annual_spend, fill = employment_title_wrapped)) +
    geom_col(alpha = 0.9,
        width = 0.8) +
  
    scale_fill_manual(values = c("#c43d4d", 
                              "#e07653", 
                              "#2057a7",
                              "#1e87a5",
                              "#8980cb",
                              "#808080")) +
  
        geom_label_repel(data = patchwork, aes(label = scales::dollar(round(actual_annual_spend))),
             size = 3,
             position = "identity",
             min.segment.length = 0.25,
             direction = "y",
             show.legend = FALSE) +
  
    theme_tq() +
  
    labs(title = stringr::str_glue("HRH Expenditure by Top Employment Titles")) +
    theme(axis.line=element_blank(),
        axis.text.y=element_blank(),
        axis.ticks=element_blank(),
        axis.title.x=element_blank(),
        axis.title.y=element_blank(),
        legend.position    = "none",
        plot.title = element_text(size = 10),
        axis.text.x = element_text(angle = 45, vjust = 1, hjust = 1)) 
   
g3 <- g2 + g1

g3

```

## Median Annual Remuneration
This scatter plot shows the median annual remuneration (i.e. estimated annual salary) of the top 10 employment titles among USAID IPs. This graph shows the range of remuneration for each employment title where each dot represents an individual’s estimated annual remuneration. 

```{r fig5, fig.height = 8, fig.width = 10, echo = FALSE, eval = TRUE}

# Prepare the remuneration df to be used
remuneration <- HRH_clean %>%
  filter(fiscal_year == 2024) %>%
  filter(funding_agency == "USAID") %>%
  filter(operating_unit == params$operating_unit) %>%
  select(employment_title, equivalent_annual_salary)

# Remove any extreme outliers, since we do not want to visualize them in our box plot
detect_outlier <- function(x) {
  # calculate first quantile
  Quantile1 <- quantile(x, probs=.01)
  # calculate third quantile
  Quantile3 <- quantile(x, probs=.98)
  # calculate inter quartile range
  IQR = Quantile3-Quantile1
  # return true or false
  x > Quantile3 + (IQR*1.5) | x < Quantile1 - (IQR*1.5)
}
 
# create remove outlier function
remove_outlier <- function(dataframe,
                            columns=names(dataframe)) {
  # for loop to traverse in columns vector
  for (col in columns) {
    # remove observation if it satisfies outlier function
    dataframe <- dataframe[!detect_outlier(dataframe[[col]]), ]
  }
  
  return(dataframe)
}

# Now use the function to remove extreme outliers from our remuneration df
remuneration <- remove_outlier(remuneration, 'equivalent_annual_salary')

# Calculate the median annual remuneration by employment title, filter for top 10 titles, pull the titles, and use this as a sorting variable
sort_list <- remuneration %>%
  group_by(employment_title) %>%
  summarise(median_remuneration = median(equivalent_annual_salary, na.rm = T),
            n = n()) %>%
  arrange(desc(median_remuneration)) %>%
  filter(n > 10) %>%
  top_n(n = 10, wt = median_remuneration) %>%
  mutate(employment_title_wrapped = str_wrap(employment_title, width = 10)) %>%
  pull(employment_title)
  
# Filter our remuneration df for the top titles in our sorting list
remuneration <- remuneration %>%
  filter(employment_title %in% sort_list)

# Now create the box plot
ggplot(remuneration, 
       aes(x=reorder(employment_title, equivalent_annual_salary, FUN = median), 
           y=equivalent_annual_salary)) + 
  
  geom_jitter(aes(colour = employment_title), width = 0.3, shape = 21, color = "#002a6c") +
  
  stat_summary(fun = median, show.legend = FALSE, geom = "crossbar", color = "black", width = 0.8) +
  
  scale_y_continuous(labels = label_dollar()) +
  
  coord_flip() + 
  
  labs(x = "Employment Title", y = "Median Annual Remuneration") + 
  
  scale_fill_manual(values = rep(NA, 2)) +
  
  theme_tq() +
  
  theme(panel.background = element_rect(fill = "white"),
        axis.line=element_blank(),
        axis.ticks=element_blank(),
        axis.title.y=element_blank(),
        plot.title = element_text(size = 13),
        text = element_text(size = 16),
        legend.position    = "none")

```

## Local Partners
This shows how much of USAID’s HRH expenditure is channeled through prime local partners for each primary program area. Note that Local Partners here are defined as local organizations that were awarded as the prime partner, and not as sub-awardees/subrecipients. If a local organization is a subrecipient while the prime partner is an international organization for a given mechanism, then the subrecipient's staffing expenditures will be considered as non-local. This should provide a picture of the level of local partner involvement in implementing PEPFAR USAID program activities prime partners.

```{r fig6, fig.height = 5, fig.width = 6, echo = FALSE, eval = TRUE}

# Create the summary df for local partners
local_partner <- HRH_clean %>%
  filter(fiscal_year == 2024) %>%
  filter(funding_agency == "USAID") %>%
  filter(operating_unit == params$operating_unit) %>%
  select(fiscal_year, funding_agency, program, is_indigenous_prime_partner, actual_annual_spend) %>%
  mutate(is_indigenous_prime_partner = case_when(is_indigenous_prime_partner == "Y" ~"Yes",
                                                 is_indigenous_prime_partner == "N" ~ "No",
                                                 TRUE ~ is_indigenous_prime_partner)) %>%
  group_by(program, is_indigenous_prime_partner) %>%
  summarise(actual_annual_spend = sum(actual_annual_spend, na.rm = T)) %>%
  arrange(desc(actual_annual_spend)) 

# Now plot the summary df
  ggplot(local_partner, aes(x = reorder(program, -actual_annual_spend), 
                          y = actual_annual_spend, 
                          fill = is_indigenous_prime_partner,
                          label = actual_annual_spend)) +
  
  geom_bar(stat = "identity", width = 0.8) +
    
  geom_label_repel(data = local_partner, aes(label = scales::dollar(round(actual_annual_spend))),
             size = 2.5,
             position = position_stack(vjust = 0.5),
             direction = "y",
             show.legend = FALSE) +
  
  xlab("Primary Program Area") +
    
  ylab("Total HRH Expenditure") +
    
  labs(fill = "Prime Local Partner?") +
    
  scale_y_continuous(labels = label_dollar()) +
    
  scale_fill_manual(values = c("Local" = "#1e87a5",
                              "International" = "#808080")) +

  theme(panel.background = element_rect(fill = "white"),
        axis.line=element_blank(),
        axis.text.y=element_blank(),
        axis.ticks=element_blank(),
        axis.title.x=element_blank(),
        axis.title.y=element_blank(),
        plot.title = element_text(size = 10)) +
  
  theme_tq() 
  
```

\pagebreak

## Work Location

This shows the work location of USAID staff in terms of HRH expenditure and individual count. Examine to identify where most staff are located and if this is reflective of the program.  

Above site staff are defined as personnel that are not directly interacting or providing services to beneficiaries (e.g. TA personnel or support staff based mostly in offices). Roving staff are direct service delivery staff that are providing services in multiple facilities.


```{r fig7, fig.height = 6, fig.width = 8, echo = FALSE, eval = TRUE}

# Create a df that shows HRH expenditure and individual count by work location
patchwork <- HRH_clean %>%
  filter(fiscal_year == 2024) %>%
  filter(funding_agency == "USAID") %>%
  filter(operating_unit == params$operating_unit) %>%
  select(fiscal_year, work_location, actual_annual_spend, individual_count) %>%
  group_by(work_location) %>%
  summarise(actual_annual_spend = sum(actual_annual_spend, na.rm = T),
            individual_count = sum(individual_count, na.rm = T)) %>%
  arrange(desc(individual_count)) 

# Make the first plot
g1 <- ggplot(data = patchwork, mapping = aes(x = reorder(work_location, -actual_annual_spend), y = individual_count, fill = work_location)) +
    geom_col(alpha = 0.9,
        width = 0.8) +
  
    scale_fill_manual(values = c("Above Site" = "#c43d4d", 
                              "Facility" = "#2057a7",
                              "Community" = "#1e87a5",
                              "Roving" = "#8980cb")) +
  
    geom_label_repel(data = patchwork, aes(label = scales::comma(round(individual_count))),
             size = 4,
             position = "identity",
             min.segment.length = 0.25,
             direction = "y",
             show.legend = FALSE) +
  
  
    theme_tq() +
  
    labs(title = stringr::str_glue("Individual Count by Work Location")) +
    theme(axis.line=element_blank(),
        axis.text.y=element_blank(),
        axis.ticks=element_blank(),
        axis.title.x=element_blank(),
        axis.title.y=element_blank(),
        legend.position    = "none",
        plot.title = element_text(size = 14),
        text = element_text(size = 16)) 

# Make the second plot

g2 <- ggplot(data = patchwork, mapping = aes(x = reorder(work_location, -actual_annual_spend), y = actual_annual_spend, fill = work_location)) +
    geom_col(alpha = 0.9,
        width = 0.8) +
  
    scale_fill_manual(values = c("Above Site" = "#c43d4d", 
                              "Facility" = "#2057a7",
                              "Community" = "#1e87a5",
                              "Roving" = "#8980cb")) +
  
    geom_label_repel(data = patchwork, aes(label = scales::dollar(round(actual_annual_spend))),
             size = 4,
             position = "identity",
             min.segment.length = 0.25,
             direction = "y",
             show.legend = FALSE) +
  
    theme_tq() +
  
    labs(title = stringr::str_glue("HRH Expenditure by Work Location")) +
    theme(axis.line=element_blank(),
        axis.text.y=element_blank(),
        axis.ticks=element_blank(),
        axis.title.x=element_blank(),
        axis.title.y=element_blank(),
        legend.position    = "none",
        plot.title = element_text(size = 14),
        text = element_text(size = 16)) 
   
g3 <- g2 + g1

g3

```

\pagebreak

## Top Employment Titles for Community-Based Staff
This shows the most common employment title for community-based staff by count

```{r fig8, fig.height = 5, fig.width = 8, echo = FALSE, eval = TRUE}

# Create a df that shows HRH expenditure and individual count by the top 5 community titles
patchwork <- HRH_clean %>%
  filter(fiscal_year == 2024) %>%
  filter(funding_agency == "USAID") %>%
  filter(operating_unit == params$operating_unit) %>%
  filter(is_community_primarily == "Yes") %>%
  select(fiscal_year, work_location, employment_title, actual_annual_spend, individual_count) %>%
  group_by(employment_title) %>%
  summarise(actual_annual_spend = sum(actual_annual_spend, na.rm = T),
            individual_count = sum(individual_count, na.rm = T)) %>%
  top_n(n = 5, wt = individual_count) %>%
  ungroup() %>%
  mutate(employment_title_wrapped = str_wrap(employment_title, width = 10)) %>%
  select(-employment_title)

# Make the column graph
g1 <- patchwork %>%
    ggplot(aes(x = reorder(employment_title_wrapped, individual_count))) +
    geom_col(
        aes(y = individual_count, fill = "Individual Count"),
        alpha = 0.9,
        width = 0.8
    ) +
    geom_label(
        aes(
            y     = individual_count,
            label = scales::comma(round(individual_count)),
            color = "Individual Count",
        ),
        size = 4
    ) +
  
  ylab("Individual Count") + 
  
  theme_tq() +
  
  theme(axis.line=element_blank(),
        axis.ticks=element_blank(),
        plot.title = element_text(size = 10),
        text = element_text(size = 14)) +
  
    scale_color_manual(values = c(
        palette_light()[["blue"]]
    )) +
  
    scale_fill_manual(values = palette_light()[["blue"]]) +
    theme(
        legend.position    = "none",
        axis.title.y=element_blank()
    ) +

    scale_y_continuous(labels = scales::comma_format()) +
  
    coord_flip()

g1

```

## Top Primary Beneficiaries for Community-Based Staff
This shows the top primary beneficiaries supported by community-based staff by count

```{r fig9, fig.height = 5, fig.width = 8, echo = FALSE, eval = TRUE}

# Create a df that shows HRH expenditure and individual count by primary beneficiary
patchwork <- HRH_clean %>%
  filter(fiscal_year == 2024) %>%
  filter(funding_agency == "USAID") %>%
  filter(operating_unit == params$operating_unit) %>%
  filter(work_location == "Community") %>%
  select(fiscal_year, targeted_beneficiary, actual_annual_spend, individual_count) %>%
  mutate(beneficiary_wrapped = str_wrap(targeted_beneficiary, width = 10)) %>%
  group_by(beneficiary_wrapped) %>%
  summarise(actual_annual_spend = sum(actual_annual_spend, na.rm = T),
            individual_count = sum(individual_count, na.rm = T)) %>%
  arrange(desc(individual_count)) 

# Make the first plot
g1 <- ggplot(data = patchwork, mapping = aes(x = reorder(beneficiary_wrapped, -actual_annual_spend), y = individual_count, fill = beneficiary_wrapped)) +
    geom_col(alpha = 0.9,
        width = 0.8) +
  
    scale_fill_manual(values = c("Military" = "#c43d4d", 
                              "Key\nPopulations" = "#e07653", 
                              "Children" = "#2057a7",
                              "OVC" = "#1e87a5",
                              "AGYW" = "#f2bc40",
                              "Pregnant & Breastfeeding Women" = "#8980cb",
                              "Non-Targeted\nPopulations" = "#808080")) +
  
    geom_label_repel(data = patchwork, aes(label = scales::comma(round(individual_count))),
             size = 3.5,
             position = "identity",
             min.segment.length = 0.25,
             direction = "y",
             show.legend = FALSE) +
  
    theme_tq() +
  
    labs(title = stringr::str_glue("Individual Count by Primary Beneficiary")) +
    theme(axis.line=element_blank(),
        axis.text.y=element_blank(),
        axis.ticks=element_blank(),
        axis.title.x=element_blank(),
        axis.title.y=element_blank(),
        legend.position    = "none",
        plot.title = element_text(size = 12),
        text = element_text(size = 13),
        axis.text.x = element_text(angle = 45, vjust = 1, hjust = 1)) 


# Make the second plot

g2 <- ggplot(data = patchwork, mapping = aes(x = reorder(beneficiary_wrapped, -actual_annual_spend), y = actual_annual_spend, fill = beneficiary_wrapped)) +
    geom_col(alpha = 0.9,
        width = 0.7) +
  
    scale_fill_manual(values = c("Military" = "#c43d4d", 
                              "Key\nPopulations" = "#e07653", 
                              "Children" = "#2057a7",
                              "OVC" = "#1e87a5",
                              "AGYW" = "#f2bc40",
                              "Pregnant & Breastfeeding Women" = "#8980cb",
                              "Non-Targeted\nPopulations" = "#808080")) +
  
    geom_label_repel(data = patchwork, aes(label = scales::dollar(round(actual_annual_spend))),
             size = 3.5,
             position = "identity",
             min.segment.length = 0.25,
             direction = "y",
             show.legend = FALSE) +
  
    theme_tq() +
  
    labs(title = stringr::str_glue("HRH Expenditure by Primary Beneficiary")) +
    theme(axis.line=element_blank(),
        axis.text.y=element_blank(),
        axis.ticks=element_blank(),
        axis.title.x=element_blank(),
        axis.title.y=element_blank(),
        legend.position    = "none",
        plot.title = element_text(size = 12),
        text = element_text(size = 13),
        axis.text.x = element_text(angle = 45, vjust = 1, hjust = 1)) 
   
g3 <- g2 + g1

g3

```

## Top Prime Implementation Partners 

This shows the top 5 prime partners with the highest staffing footprint in terms of HRH expenditure. Each prime partner is disaggregated by primary program area to indicate the program activities that each partner is primarily engaged in. If requested, we can also provide a deeper dive review of each top implementation partner to better describe their staffing composition.


```{r fig10, fig.height = 6, fig.width = 8, echo = FALSE, eval = TRUE}

# Pull names of top 5 partners only
top_5_partners <- HRH_clean %>%
  filter(fiscal_year == 2024) %>%
  filter(funding_agency == "USAID") %>%
  filter(operating_unit == params$operating_unit) %>%
  group_by(prime_partner_name) %>%
  summarise(actual_annual_spend = sum(actual_annual_spend, na.rm = T)) %>%
  arrange(desc(actual_annual_spend)) %>%
  top_n(n = 5, wt = actual_annual_spend) %>%
  pull(prime_partner_name)

# Create a summary table that filters for the top 5 partner names only
prime_partners <- HRH_clean %>%
  filter(fiscal_year == 2024) %>%
  filter(funding_agency == "USAID") %>%
  filter(operating_unit == params$operating_unit) %>%
  filter(prime_partner_name %in% top_5_partners) %>%
  select(fiscal_year, er_category, prime_partner_name, actual_annual_spend) %>%
  group_by(prime_partner_name, er_category) %>%
  summarise(actual_annual_spend = sum(actual_annual_spend, na.rm = T)) %>%
  arrange(desc(actual_annual_spend)) %>%
  mutate(prime_partner_name_wrapped = str_wrap(prime_partner_name, width = 10))
  
# Now make the plot
  ggplot(prime_partners, aes(x = reorder(prime_partner_name_wrapped, -actual_annual_spend), 
                             y = actual_annual_spend, 
                             fill = er_category,
                             label = actual_annual_spend)) +
  
  geom_bar(stat = "identity", width = 0.7) +
    
  geom_label_repel(data = prime_partners, aes(label=scales::dollar(round(actual_annual_spend))),
             size = 3,
             position = position_stack(vjust = 0.5),
             min.segment.length = 0.25,
          #   direction = "x",
             show.legend = FALSE) +
 
  scale_y_continuous(labels = label_dollar()) +
    
  xlab("Prime partner name") +
    
  ylab("Individual Count") + 
    
  labs(fill = "Worker Type (ER Category)") +
  
  scale_fill_manual(values = c("HCW: Ancillary" = "#287c6f", 
                              "HCW: Clinical" =  "#2057a7", 
                              "Other Staff" = "#808080", 
                              "Program Management" = "#C43d4d")) +
    
  theme_tq() +
  
  theme(axis.line=element_blank(),
        axis.text.y=element_blank(),
        axis.ticks=element_blank(),
        axis.title.x=element_blank(),
        axis.title.y=element_blank(),
        plot.title = element_text(size = 10),
        text = element_text(size = 13),
        axis.text.x = element_text(angle = 45, vjust = 1, hjust = 1)) 

```

For more detailed analysis of the `r params$fiscal_year` HRH Inventory dataset, please contact the HRH Reporting Help Desk (hrh-reporting-helpdesk@usaid.gov). This may include program-specific deep dives of C&T, DREAMS, OVC, and/or KP mechanisms or reviewing the data using more customized visuals.

